### Compute County Min, Median, Max ###

With the county division subsets now in the HDFS we will illustate using them
to carry out D&amp;R with a very simple recombination procedure based on a
summary statistic for each county of the variable `listing`.
We do this for simplicity of explanation of how `RHIPE` works.
However, we emphasize that in practice initial analysis would
almost always involve comprehensive analysis of both the detailed data for all
subset variables and summary statistics based on the detailed data.

Our summary statistic consists of the minimum, median, and maximum of
`listing`, one summary for each county. Map R code computes the statistic.
The output key of Map, and therefore the input key for Reduce is `state`.
The Reduce R code creates a `data.frame` for each state
where the columns are `FIPS`, `county`, `min`, `median`, and `max`.

The next step is to analyze the county summary statistic, so ths
is an example of an analytic recombination. In addition, we suppose that the
summary statistic dataset is small enough to analyze in R. This is not uncommon
in practice even when the raw data are very large and complex.

### The RHIPE Manager: rhwatch() ####

Here is the code for `rhwatch()`.

```{r eval=FALSE, tidy=FALSE}
CountyStats <- rhwatch(
  map      = map2,
  reduce   = reduce2,
  input    = rhfmt("/yourloginname/housing/byCounty", type = "sequence"),
  output   = rhfmt("/yourloginname/housing/CountyStats", type = "sequence"),
  readback = TRUE
  mapred   = list( 
    mapred.reduce.tasks = 10
  ),
)
```
Our Map and Reduce code, `map2` and `reduce2`, is given to the arguments
`map` and `reduce`. The code will be will be discussed later.

The initial input key-value pairs for Map, given to the argument `input`,
are our county subsets which were written to the HDFS directory
/yourloginname/housing  as the key-value pairs `list` object `byCounty`. 
The final output key-value pairs for Reduce, specified by the argument
`output`, will be written to the file `CountyStats` in the same directory as
the subsets.
The output will be a key-value pair `list` object.
The keys are the states, and the values will be a `data.frame` object for each
observation of `state`. The variables of the `data.frame` are given above.

The argument `readback` is given the value TRUE, which means `CountyStats` is
also written to the R global environment of the R session. We do this since we
are have supposed the Reduce outputs small enough to be analyzed in R.

The argument `mapred.reduce.tasks` is given the value 10, as in
our use of it to create the county subsets.


#### The Map R Code ####

The Map R code is
```{r eval=FALSE, tidy=FALSE}
map2 <- expression({
  lapply(seq_along(map.keys), function(r) {
    outputvalue <- data.frame(
      FIPS = map.keys[[r]],
      county = attr(map.values[[r]], "county"),
      min = min(map.values[[r]]$listing, na.rm = TRUE),
      median = median(map.values[[r]]$listing, na.rm = TRUE),
      max = max(map.values[[r]]$listing, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
    outputkey <- attr(map.values[[r]], "state")
    rhcollect(outputkey, outputvalue)
  })
})
```

`map.keys` are the Map input keys, the county subset identiers `FIPS`, and
`map.values` are the Map input values,  the county subset `data.frame`
objects.  The `lapply()` loop goes through all subsets.
Each stage of the loop creates the Map output vales, `outputvalue`,
a `data.frame` with one row that has the variables
`FIPS`, `county`, `min`, `median`, and `max`.
`outputkey` is the Map output key, the value of `state` for the county.

#### The Reduce R Code ####
The Reduce R code for the `listing` summary statistic is


```{r eval=FALSE, tidy=FALSE}
reduce2 <- expression(
  pre = {
    reduceoutputvalue <- data.frame()
  },
  reduce = {
    reduceoutputvalue <- rbind(reduceoutputvalue, do.call(rbind, reduce.values))
  },
  post = {
    rhcollect(reduce.key, reduceoutputvalue)
  }
)
```

We can use almost the same reduce expression which we used previously. In the reduce expression 
above, our goal is to combine all observations associated with one particular state into a single 
`data.frame`.  In `pre`, we initialize an empty data frame, `reduceoutputvalue`. In `reduce`, we 
use `rbind` to combine all counties with their min, median, and max associated with one particular 
state.  In `post`, we just emit the final key-value pair.  The key is the state name, and the value 
is the data frame with all counties belonging to that state. These final key-value pairs are written 
to HDFS, and will persist for subsequent analyses.


This time we specify the `readback` argument to be `TRUE` in `rhwatch()` function. By doing this,
once the job is finished, the output of the job which is one `data.frame` for each state will be 
read back and assigned to `countyStats` automatically. So we do not nead to use `rhread()` anymore.

```{r eval=FALSE, tidy=FALSE}
str(CountyStats)
```
```
List of 49
 $ :List of 2
  ..$ : chr "AL"
  ..$ :'data.frame':  64 obs. of  5 variables:
  .. ..$ min   : num [1:64] 34.9 94.7 83.9 92.9 60.3 ...
  .. ..$ median: num [1:64] 51.9 99.2 88.6 97.5 72.5 ...
  .. ..$ max   : num [1:64] 73.5 102.2 94.7 105.7 93.5 ...
  .. ..$ FIPS  : chr [1:64] "01093" "01051" "01031" "01125" ...
  .. ..$ county: chr [1:64] "Marion" "Elmore" "Coffee" "Tuscaloosa" ...
 $ :List of 2
  ..$ : chr "AR"
  ..$ :'data.frame':	71 obs. of  5 variables:
  .. ..$ min   : num [1:71] 64.916 0.548 58.333 31.409 92.222 ...
  .. ..$ median: num [1:71] 84.3 59.6 68.4 40 95.9 ...
  .. ..$ max   : num [1:71] 97.9 222 75.9 46.2 105.2 ...
  .. ..$ FIPS  : chr [1:71] "05105" "05079" "05071" "05011" ...
  .. ..$ county: chr [1:71] "Perry" "Lincoln" "Johnson" "Bradley" ...
......
```

We can access the key of the first key-value pair by

```{r eval=FALSE, tidy=FALSE}
CountyStats[[1]][[2]]
```
```
[1] "AL"
```
And the `data.frame` for state "AL" is:

```{r eval=FALSE, tidy=FALSE}
head(CountyStats[[1]][[2]])
```
```
        min    median       max  FIPS     county
1  34.88372  51.88628  73.46257 01093     Marion
2  94.66667  99.20582 102.23077 01051     Elmore
3  83.93817  88.59977  94.67041 01031     Coffee
4  92.87617  97.53306 105.71429 01125 Tuscaloosa
5  60.34774  72.46377  93.53741 01027       Clay
6 108.97167 119.66207 128.13390 01117     Shelby
```
