### Compute County Min, Median, Max ###

With the county division subsets now in the HDFS we will illustrate using them
to carry out D&amp;R with a very simple recombination procedure based on a
summary statistic for each county of the variable `listing`.
We do this for simplicity of explanation of how `RHIPE` works.
However, we emphasize that in practice, initial analysis would
almost always involve comprehensive analysis of both the detailed data for all
subset variables and summary statistics based on the detailed data.

Our summary statistic consists of the minimum, median, and maximum of
`listing`, one summary for each county. Map R code computes the statistic.
The output key of Map, and therefore the input key for Reduce is `state`.
The Reduce R code creates a `data.frame` for each state
where the columns are `FIPS`, `county`, `min`, `median`, and `max`.
So our example illustrates a scenario where we create summary statistics, and
then analyze the results. This is an analytic recombination. In addition, we
suppose that in this scenario the summary statistic dataset is small enough to
analyze in the standard serial R.  This is not uncommon in practice even when
the raw data are very large and complex.

### The RHIPE Manager: rhwatch() ####

Here is the code for `rhwatch()`.

```{r eval=FALSE, tidy=FALSE}
CountyStats <- rhwatch(
  map      = map2,
  reduce   = reduce2,
  input    = rhfmt("/yourloginname/housing/byCounty", type = "sequence"),
  output   = rhfmt("/yourloginname/housing/CountyStats", type = "sequence"),
  readback = TRUE
)
```
Our Map and Reduce code, `map2` and `reduce2`, is given to the arguments
`map` and `reduce`. The code will be will be discussed later.

The input key-value pairs for Map, given to the argument `input`,
are our county subsets which were written to the HDFS directory
`/yourloginname/housing`  as the key-value pairs `list` object `byCounty`. 
The final output key-value pairs for Reduce, specified by the argument
`output`, will be written to the `list` object `CountyStats` in the same
directory as the subsets. The keys are the states, and the values are the
`data.frame` objects for the states.

The argument `readback` is given the value TRUE, which means `CountyStats` is
also written to the R global environment of the R session. We do this because
our scenario is that analytic recombination is done in R.

The argument `mapred.reduce.tasks` is given the value 10, as in our use of it
to create the county subsets.


#### The Map R Code ####

The Map R code is
```{r eval=FALSE, tidy=FALSE}
map2 <- expression({
  lapply(seq_along(map.keys), function(r) {
    outputvalue <- data.frame(
      FIPS = map.keys[[r]],
      county = attr(map.values[[r]], "county"),
      min = min(map.values[[r]]$listing, na.rm = TRUE),
      median = median(map.values[[r]]$listing, na.rm = TRUE),
      max = max(map.values[[r]]$listing, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
    outputkey <- attr(map.values[[r]], "state")
    rhcollect(outputkey, outputvalue)
  })
})
```

`map.keys` is the Map input keys, the county subset identifiers `FIPS`.
`map.values` is the Map input values, the county subset `data.frame`
objects. The `lapply()` loop goes through all subsets, and the looping
variable is `r`. Each stage of the loop creates one output key-value pair,
`outputkey` and `outputvalue`.
`outputkey` is the observation of `state`. `outputvalue` is a `data.frame`
with one row that has the variables `FIPS`, `county`, `min`, `median`, and
`max` for county `FIPS`. `rhcollect(outputkey, outputvalue)` emits the pairs
to reducers, becoming the Reduce input key-value pairs.

#### The Reduce R Code ####
The Reduce R code for the `listing` summary statistic is


```{r eval=FALSE, tidy=FALSE}
reduce2 <- expression(
  pre = {
    reduceoutputvalue <- data.frame()
  },
  reduce = {
    reduceoutputvalue <- rbind(reduceoutputvalue, do.call(rbind, reduce.values))
  },
  post = {
    rhcollect(reduce.key, reduceoutputvalue)
  }
)
```

The first task of Reduce is to group its input key-value pairs by unique key,
in this case by `state`. The Reduce R code is applied to the key-value pairs
of each group by a reducer.

Expression `pre`, initializes `reduceoutputvalue` to a
`data.frame()`. `reduce` assembles the state `data.frame` as the
reducer receives the values through `rbind(reduceoutputvalue, do.call(rbind,
reduce.values))`; this uses `rbind()` to add rows to the `data.frame` object.
`post` operates further on the result of `reduce`; `rhcollect()` forms a Reduce
output key-value pair for each state. RHIPE then writes the Reduce output
key-value pairs to the HDFS.

Recall that we told RHIPE in `rhwatch()` to also write the Reduce output
to `CountyStats` in both the R server global environment. There, we can have a
look at the results to make sure all is well. We can look at a summary

```{r eval=FALSE, tidy=FALSE}
str(CountyStats)
```
```
List of 49
 $ :List of 2
  ..$ : chr "AL"
  ..$ :'data.frame':  64 obs. of  5 variables:
  .. ..$ min   : num [1:64] 34.9 94.7 83.9 92.9 60.3 ...
  .. ..$ median: num [1:64] 51.9 99.2 88.6 97.5 72.5 ...
  .. ..$ max   : num [1:64] 73.5 102.2 94.7 105.7 93.5 ...
  .. ..$ FIPS  : chr [1:64] "01093" "01051" "01031" "01125" ...
  .. ..$ county: chr [1:64] "Marion" "Elmore" "Coffee" "Tuscaloosa" ...
 $ :List of 2
  ..$ : chr "AR"
  ..$ :'data.frame':	71 obs. of  5 variables:
  .. ..$ min   : num [1:71] 64.916 0.548 58.333 31.409 92.222 ...
  .. ..$ median: num [1:71] 84.3 59.6 68.4 40 95.9 ...
  .. ..$ max   : num [1:71] 97.9 222 75.9 46.2 105.2 ...
  .. ..$ FIPS  : chr [1:71] "05105" "05079" "05071" "05011" ...
  .. ..$ county: chr [1:71] "Perry" "Lincoln" "Johnson" "Bradley" ...
......
```

We can look at the first key-value pair


```{r eval=FALSE, tidy=FALSE}
CountyStats[[1]][[2]]
```
```
[1] "AL"
```

We can look at the `data.frame` for state "AL"

```{r eval=FALSE, tidy=FALSE}
head(CountyStats[[1]][[2]])
```
```
        min    median       max  FIPS     county
1  34.88372  51.88628  73.46257 01093     Marion
2  94.66667  99.20582 102.23077 01051     Elmore
3  83.93817  88.59977  94.67041 01031     Coffee
4  92.87617  97.53306 105.71429 01125 Tuscaloosa
5  60.34774  72.46377  93.53741 01027       Clay
6 108.97167 119.66207 128.13390 01117     Shelby
```
