### Compute County Means ###

We calculate the mean list price and mean sale price by county exactly the same way we 
calculated them by state. 

#### Map ####

```{r eval=FALSE, tidy=FALSE}
map4 <- expression({
  lapply(seq_along(map.keys), function(r) {
    value <- data.frame(
      listMean = mean(map.values[[r]]$list, na.rm = TRUE),
      saleMean = mean(map.values[[r]]$selling, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
    value$state <- attr(map.values[[r]], "state")
    value$county <- attr(map.values[[r]], "county")
    value$FIPS <- map.keys[[r]]
    rhcollect(1, value)
  })
})
```

In the map expression, we create a single row data frame for each county. The data frame has five
columns: FIPS, listMean, saleMean, state, and county name. FIPS code, state name, and county name 
can be found in the attributes of each element of `map.values`. In order to combine all means by
county into one data frame, we will assign 1 to be the key for all intermediate key-value pairs.

#### Reduce ####

```{r eval=FALSE, tidy=FALSE}
reduce4 <- expression(
  pre = {
    countyMeans <- data.frame()
  },
  reduce = {
    countyMeans <- rbind(countyMeans, do.call(rbind, reduce.values))
  },
  post = {
    rhcollect(reduce.key, countyMeans)
  }
)
```

We can use the same reduce expression which we used to find the means by state. The final output
consists of one key-value pair, where the key is 1, and value is the data frame with all county
means.

#### Execution Function ####

```{r eval=FALSE, tidy=FALSE}
meansByCounty <- rhwatch(
  map      = map4,
  reduce   = reduce4,
  input    = rhfmt("/ln/tongx/housing/byCounty", type = "sequence"),
  output   = rhfmt("/ln/tongx/housing/meansByCounty", type = "sequence"),
  mapred   = list( 
    mapred.reduce.tasks = 1
  ),
  readback = TRUE
)
```

This time we specify only one reduce task by setting `mapred.reduce.tasks` to be 1.
Since we know there is only one
key-value pair in the output, we only need one output file. Eliminating unnecessary output files 
can speed up this job and future jobs which read its output.

```{r eval=FALSE, tidy=FALSE}
str(meansByCounty)
```
```
List of 1
 $ :List of 2
  ..$ : num 1
  ..$ :'data.frame':  2883 obs. of  5 variables:
  .. ..$ listMean: num [1:2883] 66.3 74.9 86.5 68.3 74.4 ...
  .. ..$ saleMean: num [1:2883] NaN NaN 66.6 52.2 NaN ...
  .. ..$ state   : chr [1:2883] "AL" "AL" "AL" "AL" ...
  .. ..$ county  : chr [1:2883] "Chambers" "Colbert" "Cullman" "Lawrence" ...
  .. ..$ FIPS    : chr [1:2883] "01017" "01033" "01043" "01079" ...
```
```{r eval=FALSE, tidy=FALSE}
head(meansByCounty[[1]][[2]])
```
```
   listMean saleMean state   county  FIPS
1  66.26665      NaN    AL Chambers 01017
2  74.91187      NaN    AL  Colbert 01033
3  86.54870 66.56352    AL  Cullman 01043
4  68.32503 52.17811    AL Lawrence 01079
5  74.35722      NaN    AL    Macon 01087
6 103.70687 94.27969    AZ   Mohave 04015
```
