### Compute County Min, Median, Max ###

With the county division subsets now in the HDFS we will illustate using them
to carry out D&amp;R with a very simple recombination procedure for a summary
statistic of the variable `listing` for each county, but one that suffices to
illustrate
`RHIPE`.  However, we emphasize that in practice initial analysis would
actually involve comprehensive analysis of both the detailed data for all
subset variables and summary statistics based on the detailed data.

The summary statistics consists of the minimum, median, and maximum of
`listing` one for each county. Map R code computes the statistic.
The output key of Map, and the therefore the input key for Reduce is `state`.
The Reduce R code creates a `data.frame` for each state
where the columns are `min`, `median`, `max`, and `county`.

We suppose that we in a situation in which while the dataset is large and
complex, requiring RHIPE, but that the summary statistic data is small enought
to analyze further in R. This is an example of an analytic recombintion.

### The RHIPE Manager: rhwatch() ####

Here is the code for `rhwatch()`. Our Map and Reduce code,
`map2` and `reduce2` will be discussed later.

The initial input key-value pairs for Map, given to the argument `input`,
are our county subsets which were written to the HDFS directory
/yourloginname/housing  as the key-value pairs `list` object `byCounty`. 
The final output key-value pairs for Reduce, specified by the argument
`output`, will be in the same directory in the file CountyStats.
This will be a key-value pair `list` object. The keys are the states, and the
values will be a `data.frame` object for each observation of `state`. The
variabls are the county `listing` summary statistic variables,
the observations of `FIPS`, and the observation of `county`.

```{r eval=FALSE, tidy=FALSE}
CountyStats <- rhwatch(
  map      = map2,
  reduce   = reduce2,
  input    = rhfmt("/yourloginname/housing/byCounty", type = "sequence"),
  output   = rhfmt("/yourloginname/housing/CountyStats", type = "sequence"),
  readback = TRUE
  mapred   = list( 
    mapred.reduce.tasks = 10
  ),
)
```
The argument `readback` is given the value TRUE, which means `CountyStats` is
also written to the R global environment of the R session. We do this since we
are thinking of the housing data as an example where the outputs are a dataset
small enough to be analyzed further by an analytic recombination in R without
`RHIPE`. The argument `mapred.reduce.tasks` is given the value 10, as in
our last use of it.


#### The Map R Code ####

```{r eval=FALSE, tidy=FALSE}
map2 <- expression({
  lapply(seq_along(map.keys), function(r) {
    outputvalue <- data.frame(
      FIPS = map.keys[[r]],
      county = attr(map.values[[r]], "county"),
      min = min(map.values[[r]]$listing, na.rm = TRUE),
      median = median(map.values[[r]]$listing, na.rm = TRUE),
      max = max(map.values[[r]]$listing, na.rm = TRUE),
      stringsAsFactors = FALSE
    )
    outputkey <- attr(map.values[[r]], "state")
    rhcollect(outputkey, outputvalue)
  })
})
```

`map.keys` are the county identiers `FIPS`, and map.values are the county
subset `data.frame` objects.  The `lapply()` loop goes through all
observations of `FIPS`. `RHIPE` gets these from 
Each stage of the loop creates `outputvalue`,
a `data.frame` with one row that has as variables `min`, `median`, and `max`
for `listing`, together with the `FIPS` county observation. 
For this value, `outputkey` is the value of `state` for the county.


In the map expression, we create a single row data frame `outputvalue` for each county. The data frame has five
columns: FIPS code, county name, min of listing price, median of listing price, and max of listing price. 
FIPS code and county name can be found in the attributes of each element of `map.values`. Recall 
that those attributes are created in last session when we created county divisions.
In order to combine  by county into one data frame for each state, we will assign the state name
which also can be found in the attributes named "state" in each element of `map.values` to be the 
`outputkey` for all output key-value pairs.

#### The Reduce R Code ####

```{r eval=FALSE, tidy=FALSE}
reduce2 <- expression(
  pre = {
    reduceoutputvalue <- data.frame()
  },
  reduce = {
    reduceoutputvalue <- rbind(reduceoutputvalue, do.call(rbind, reduce.values))
  },
  post = {
    rhcollect(reduce.key, reduceoutputvalue)
  }
)
```

We can use almost the same reduce expression which we used previously. In the reduce expression 
above, our goal is to combine all observations associated with one particular state into a single 
`data.frame`.  In `pre`, we initialize an empty data frame, `reduceoutputvalue`. In `reduce`, we 
use `rbind` to combine all counties with their min, median, and max associated with one particular 
state.  In `post`, we just emit the final key-value pair.  The key is the state name, and the value 
is the data frame with all counties belonging to that state. These final key-value pairs are written 
to HDFS, and will persist for subsequent analyses.


This time we specify the `readback` argument to be `TRUE` in `rhwatch()` function. By doing this,
once the job is finished, the output of the job which is one `data.frame` for each state will be 
read back and assigned to `countyStats` automatically. So we do not nead to use `rhread()` anymore.

```{r eval=FALSE, tidy=FALSE}
str(CountyStats)
```
```
List of 49
 $ :List of 2
  ..$ : chr "AL"
  ..$ :'data.frame':  64 obs. of  5 variables:
  .. ..$ min   : num [1:64] 34.9 94.7 83.9 92.9 60.3 ...
  .. ..$ median: num [1:64] 51.9 99.2 88.6 97.5 72.5 ...
  .. ..$ max   : num [1:64] 73.5 102.2 94.7 105.7 93.5 ...
  .. ..$ FIPS  : chr [1:64] "01093" "01051" "01031" "01125" ...
  .. ..$ county: chr [1:64] "Marion" "Elmore" "Coffee" "Tuscaloosa" ...
 $ :List of 2
  ..$ : chr "AR"
  ..$ :'data.frame':	71 obs. of  5 variables:
  .. ..$ min   : num [1:71] 64.916 0.548 58.333 31.409 92.222 ...
  .. ..$ median: num [1:71] 84.3 59.6 68.4 40 95.9 ...
  .. ..$ max   : num [1:71] 97.9 222 75.9 46.2 105.2 ...
  .. ..$ FIPS  : chr [1:71] "05105" "05079" "05071" "05011" ...
  .. ..$ county: chr [1:71] "Perry" "Lincoln" "Johnson" "Bradley" ...
......
```

We can access the key of the first key-value pair by

```{r eval=FALSE, tidy=FALSE}
CountyStats[[1]][[2]]
```
```
[1] "AL"
```
And the `data.frame` for state "AL" is:

```{r eval=FALSE, tidy=FALSE}
head(CountyStats[[1]][[2]])
```
```
        min    median       max  FIPS     county
1  34.88372  51.88628  73.46257 01093     Marion
2  94.66667  99.20582 102.23077 01051     Elmore
3  83.93817  88.59977  94.67041 01031     Coffee
4  92.87617  97.53306 105.71429 01125 Tuscaloosa
5  60.34774  72.46377  93.53741 01027       Clay
6 108.97167 119.66207 128.13390 01117     Shelby
```
