### Access Subset by State ###

As we have already
seen in previous examples, our R objects from a mapreduce job have been saved as a sequence file on HDFS.
We also have seen that how to control the number of key-value pairs we want to read by specifying the `max`
argument in `rhread`. However, this can only guarantee we read fixed number of key-value pairs. The order
of the key-value pairs or those R objects saved on HDFS is random. What if we would like to only access 
specific key-value pairs by the key without reading all key-value pairs back into the R global environment?
In other words, can we treat our dataset as a queryable database through `Rhipe` functions? The answer is
absolutely YES!

In order to achieve this purpose, we have to save our subsets as a map file instead of sequence file on HDFS.
Map file is another type of file can be saved on HDFS besides text and sequence file. For example, we are 
still trying to get subsets for each state based on the subsets for each county. This time we would like to 
be able to access the subset of data for Indiana state only. The code is showed as following:

```{r eval=FALSE, tidy=FALSE}
map5 <- expression({
  lapply(seq_along(map.keys), function(r) {
    key <- attr(map.values[[r]], "state")
    county <- attr(map.values[[r]], "county")
    value <- map.values[[r]]
    value$fips <- rep(map.keys[[r]], nrow(value))
    value$county <- rep(county, nrow(value))
    rhcollect(key, value)
  })
})
reduce5 <- expression(
  pre = {
    combine <- data.frame()
  },
  reduce = {
    combine <- rbind(combine, do.call(rbind, reduce.values))
  },
  post = {
    rhcollect(reduce.key, combine)
  }
)
mr5 <- rhwatch(
  map      = map5,
  reduce   = reduce5,
  input    = rhfmt("/user/tongx/housing/bycounty", type = "sequence"),
  output   = rhfmt("/user/tongx/housing/countytostate.map", type = "map"),
  mapred   = list( 
    mapred.reduce.tasks = 10
  ),
  readback = FALSE
)
```

It is not hard to notice that the only difference between the code above with the code in the previous
session is that the `type` in `output` argument in `rhwatch` function is changed to be `map`, which means
we are creating a map file output in this mapreduce job. In this mapreduce job, we are doing exactly same
division of the dataset as we did in sequence file, except that we would like to add indexing by key 
property in the output file. And this is the main difference between map file and sequence file. After
the job is finished, we can access a map file with `rhmapfile` function.

```{r eval=FALSE, tidy=FALSE}
rst <- rhmapfile("/user/tongx/housing/countytostate.map")
rst
```
```
/ln/tmp/housing/countytostate.map is a MapFile with 10 index files
```
```{r eval=FALSE, tidy=FALSE}
class(rst)
```
```
[1] "mapfile"
```
```{r eval=FALSE, tidy=FALSE}
object.size(rst)
```
```
264 bytes
```

We can see that `rst` is a MapFile with 10 index files because we specified the reduce tasks number 
to be 10. Its class is `mapfile`, and size is only 264 bytes, which we did not read the whole R
object back from HDFS.

Now let us have a deeper look at the output files on HDFS. Recall that "/user/tongx/housing/countytostate"
is the sequence output files, and "/user/tongx/housing/countytostate.map" is the map output files.

```{r eval=FALSE, tidy=FALSE}
rhls("/user/tongx/housing/countytostate.map")
```
```
   permission owner      group size          modtime                                               file
1  -rw-r--r-- tongx supergroup    0 2014-09-28 13:52     /user/tongx/housing/countytostate.map/_SUCCESS
2  drwxrwxrwx tongx supergroup    0 2014-09-28 13:51        /user/tongx/housing/countytostate.map/_logs
3  drwxr-xr-x tongx supergroup    0 2014-09-28 13:52 /user/tongx/housing/countytostate.map/part-r-00000
4  drwxr-xr-x tongx supergroup    0 2014-09-28 13:52 /user/tongx/housing/countytostate.map/part-r-00001
5  drwxr-xr-x tongx supergroup    0 2014-09-28 13:52 /user/tongx/housing/countytostate.map/part-r-00002
......
```
```{r eval=FALSE, tidy=FALSE}
rhls("/ln/tmp/housing/countytostate")
```
```
   permission owner      group     size          modtime                                           file
1  -rw-r--r-- tongx supergroup        0 2014-09-20 18:02     /user/tongx/housing/countytostate/_SUCCESS
2  drwxrwxrwx tongx supergroup        0 2014-09-20 18:02        /user/tongx/housing/countytostate/_logs
3  -rw-r--r-- tongx supergroup 235.1 kb 2014-09-20 18:02 /user/tongx/housing/countytostate/part-r-00000
4  -rw-r--r-- tongx supergroup   969 kb 2014-09-20 18:02 /user/tongx/housing/countytostate/part-r-00001
5  -rw-r--r-- tongx supergroup 1.698 mb 2014-09-20 18:02 /user/tongx/housing/countytostate/part-r-00002
......
```

We found out that all `part-r-` files now becomes directory instead of data files. When we go inside of each
`part-r-` files, we notice that we have two files which are `data` and `index`. `data` file is similar with 
the output in `part-r-` files we got in sequence output files. We can actually read it by using `rhread` 
function. This extra `index` file with size around 200 bytes records all indexing information by key of this 
part file.

```{r eval=FALSE, tidy=FALSE}
rhls("/user/tongx/housing/countytostate.map/part-r-00000")
```
```
  permission owner      group        size          modtime                                                     file
1 -rw-r--r-- tongx supergroup    235.1 kb 2014-09-28 13:52  /user/tongx/housing/countytostate.map/part-r-00000/data
2 -rw-r--r-- tongx supergroup   209 bytes 2014-09-28 13:52 /user/tongx/housing/countytostate.map/part-r-00000/index
```

Finally, we can access the subset of data frame of Indiana state by:

```{r eval=FALSE, tidy=FALSE}
head(rst[["IN"]])
```
```
        time  sold      list  selling  fips county
1 2010-11-01    NA  56.70013       NA 18177  Wayne
2 2010-12-01    NA  52.01389       NA 18177  Wayne
3 2011-01-01    NA  49.92504       NA 18177  Wayne
4 2011-02-01    NA  51.91466       NA 18177  Wayne
5 2011-03-01    NA  55.27211       NA 18177  Wayne
6 2011-04-01    NA  56.68348       NA 18177  Wayne
```