### Plots by State ###

After previous several sessions, we have seen how to do divide and recombine in `RHIPE`, how to 
access the subset by key. Another useful and powerful aspect of `RHIPE` is that we can achieve
parallel plotting on cluster. Visualizing each subset can help us to have a better understanding of
the subsets as well as the whole dataset.

For each subset by state, we would like to know what the number of sold units looks like over time
conditional on county. So we will plot the sold units against time for each state. When we are 
facing a relative small data set in base R, this plotting for each state has to be done either using
a for loop or some other package like `plyr` package. But when we are facing a large data set that
cannot be fit into the memory, previous visualization method will fail easily. 

In `RHIPE`, on the other hand, we can specify multiple tasks, and in each tasks we can plot one or
multiple plots. Later on we will see, in our housing example, we will specify 12 tasks to conduct 
total 49 plots, one plot for state. So each task will handle about 4 plots which will 
dramatically decreasing the total time to finish these 49 plots comparing with doing this using for
loop.

#### Time Series Plot of Sold Units ####

The first type of plot we are going to plot is the time series plot of number of sold units condition
on county for each state. First of all, let us have a look at what the plot looks like.

[Time series plot of number of units sold for North Carolina State](./plots/time.vs.unit.NC.pdf).

[Time series plot of number of units sold for Texas State](./plots/time.vs.unit.TX.pdf).

For each state we created the time series plot of the number of sold units conditional on county.
The plotting function we are going to use is in a R package named `lattice` which is a very powerful
visualization method in R. 

```{r eval=FALSE, tidy=FALSE}
map11 <- expression({
  lapply(seq_along(map.keys), function(r) {
    value <- subset(map.values[[r]], !is.na(as.numeric(units)))
    if(nrow(value) != 0) {
      rhcollect(map.keys[[r]], value)
    }
  })
})
```

In map expression, we read in each subset by state, and then we only collected subsets that is not all `NA`
for `units` variable. This can help us illuminate the number of key-value pairs that are transferred to 
reducer.

```{r eval=FALSE, tidy=FALSE}
ylab <- "Log of Number of Sold Units(log base 2)"
xlab <- "Month"
reduce11 <- expression(
  pre = {
    onestate <- data.frame()
  },
  reduce = {
    onestate <- do.call(rbind, reduce.values)
  },
  post = {
    trellis.device(pdf, 
      file  = paste("./tmp/time.vs.unit", reduce.key, "pdf", sep="."), 
      color = TRUE,
      paper = "legal"
    )
    b <- xyplot(
      log2(as.numeric(units)) ~ as.numeric(date) | FIPS,
      data   = onestate,
      layout = c(4, 3),
      cex    = 0.5,
      pch    = 16,
      scales = list(
        x = list(tick.number = 5),
        y = list(relation = "sliced") 
      ),
      xlab   = xlab,
      ylab   = ylab,
      main   = reduce.key,
    )
    print(b)
    dev.off()    
  }
)
```

In reduce expression, we did the actual plotting task. One thing should be noticed is that we defined two
string variables `xlab` and `ylab` in our initiating R global environment. We will use this two string 
variables as the labels for x axis and y axis respectively. Later on we will show how to let every
task be aware of these two objects. In the `post`, we called the `xyplot` function in `lattice` package to 
plot the `units` against `date`, and created a pdf plot file under the location or path of `./tmp/`. 

What is this `./tmp/` directory? It is a temporary working directory on each reducer. Recall that on the 
initiating R server(front-end), we have a concept of R local working directory. This temporary working 
directory is the same concept but on each sever of the Hadoop severs(back-end). Why we want to save the plots 
to this specific directory on each hadoop sever? Because `RHIPE` can copy all files in this particular 
directory to HDFS, then we can copy those files from HDFS to our R local working directory on initiating R 
sever. Finally, you can choose to view those files either through the Internet or copy them to your laptop to 
view them.

```{r eval=FALSE, tidy=FALSE}
mr11 <- rhwatch(
  map        = map11,
  reduce     = reduce11,
  input      = rhfmt("/ln/tongx/housing/byState", type = "sequence"),
  output     = rhfmt("/ln/tongx/housing/timeplotbystate", type = "sequence"),
  setup      = expression(
    map = {
    },
    reduce   = {
      library(lattice)
    }
  ),
  parameters = list(
    xlab = xlab,
    ylab = ylab
  ),
  mapred     = list( 
    mapred.reduce.tasks = 12
  ),
  copyFiles  = TRUE,
  readback   = FALSE
)
```

In the execution function, we saw a group of new arguments shown up. Let's go through them one by one. First 
is `setup` argument. It is an expression of R code to be run before map and reduce. Here the `map` in `setup`
is empty, but we called `lattice` library in `reduce`, because we would like to use the `xyplot` function in
reduce expression. `setup` is a very good place for us to set up for map and reduce respectively,like calling
library.

Second is `parameters` argument. This argument is how we distributed the `xlab` and `ylab` to every map and 
reduce task. Basically we group all objects as a list which we would like to use in each task. Then in map
or reduce expression, we can just use these objects in `parameters` as they have been already defined.

The last one is `copyFiles`. As we discussed before, when `copyFiles` is `TRUE`, all files created in 
`./tmp/` on each Hadoop severs will be copied to the output directory on HDFS.

Once the job is finished, let's see what we have in the output directory on HDFS.

```{r eval=FALSE, tidy=FALSE}
rhls("/ln/tongx/housing/timeplotbystate")
```
```
   permission owner      group        size          modtime                                           file
1  -rw-r--r-- tongx supergroup           0 2014-10-02 21:20     /ln/tongx/housing/timeplotbystate/_SUCCESS
2  drwxrwxrwx tongx supergroup           0 2014-10-02 21:20        /ln/tongx/housing/timeplotbystate/_logs
3  drwxr-xr-x tongx supergroup           0 2014-10-02 21:20     /ln/tongx/housing/timeplotbystate/_outputs
4  -rw-r--r-- tongx supergroup    94 bytes 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/part-r-00000
5  -rw-r--r-- tongx supergroup    94 bytes 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/part-r-00001
......
```

Different than all the results we had before, here we got an extra directory named `_outputs` in output directory.
When we went into it, we found following information:

```{r eval=FALSE, tidy=FALSE}
rhls("/ln/tongx/housing/timeplotbystate/_outputs")
```
```
   permission owner      group     size          modtime                                                           file
1  -rw-r--r-- tongx supergroup 53.21 kb 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.AL.pdf
2  -rw-r--r-- tongx supergroup 136.6 kb 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.AR.pdf
3  -rw-r--r-- tongx supergroup 31.86 kb 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.AZ.pdf
4  -rw-r--r-- tongx supergroup 150.6 kb 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.CA.pdf
5  -rw-r--r-- tongx supergroup 106.7 kb 2014-10-02 21:20 /ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.CO.pdf
......
```

Yes, all the pdf plots are saved here, on HDFS! The last step would be copy those pdf files from HDFS to our R local
working directory on initiating R server.

```{r eval=FALSE, tidy=FALSE}
rhget("/ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.NC.pdf", "./")
rhget("/ln/tongx/housing/timeplotbystate/_outputs/time.vs.unit.TX.pdf", "./")
```

The function that used to copy from HDFS to R local working directory is `rhget`, which is the opposite operation as 
`rhput` we used in previous session. Here we only showed copying two of those pdf files.

#### Time Series Plot of List Price ####

The second type of plot we are going to plot is the time series plot of list price condition on county for each state.
Again, let us have a look at what the plot looks like.

[Time series plot of list price for California State](./plots/list.vs.time.CA.pdf).

[Time series plot of list price for Virginia State](./plots/list.vs.time.VA.pdf).

```{r eval=FALSE, tidy=FALSE}
map12 <- expression({
  lapply(seq_along(map.keys), function(r) {
    value <- subset(
      map.values[[r]], !is.na(as.numeric(list))
    )
    if(nrow(value) != 0) {
      rhcollect(map.keys[[r]], value)
    }
  })
})
```

Same as before, we read in each subset by state in map expression, and then we only collected subsets that is not all 
`NA` for `list` variable.

```{r eval=FALSE, tidy=FALSE}
ylab <- "Log of List Price (log base 2 dollars per square foot)"
xlab <- "Month"
rhsave(
    list = c("ylab", "xlab"), 
    file = "/ln/tongx/housing/shared/label.RData"
)
```

Before we start the reduce expression, we still defined `xlab` and `ylab` correspondingly. But this time we save these
two objects as an `.RData` unto HDFS. The `rhsave` function writes an external representation of R objects to the 
specified file as `.RData` unto HDFS. Later on we will explain with more details about the reason for this.

```{r eval=FALSE, tidy=FALSE}
reduce12 <- expression(
  pre = {
    onestate <- data.frame()
  },
  reduce = {
    onestate <- do.call(rbind, reduce.values)
  },
  post = {
    trellis.device(pdf,
      file  = paste("./tmp/list.vs.time", reduce.key, "pdf", sep="."), 
      color = TRUE,
      paper = "legal"
    )
    b <- xyplot(
      log2(as.numeric(list)) ~ as.numeric(date) | FIPS,
      data   = onestate,
      layout = c(4,3),
      cex    = 0.5,
      pch    = 16,
      scales = list(
        x = list(tick.number = 7),
        y = list(relation = "sliced") 
      ),
      xlab   = xlab,
      ylab   = ylab,
      main   = reduce.key, 
    )
    print(b)
    dev.off()    
  }
)
```

Very similar reduce expression, we did the actual plotting task. 

```{r eval=FALSE, tidy=FALSE}
mr12 <- rhwatch(
  map       = map12,
  reduce    = reduce12,
  input     = rhfmt("/ln/tongx/housing/byState", type = "sequence"),
  output    = rhfmt("/ln/tongx/housing/listplotbystate", type = "sequence"),
  setup     = expression(
    reduce = {
      library(lattice)
      load("label.RData")
    }
  ),
  shared    = c("/ln/tongx/housing/shared/label.RData"),
  mapred    = list( 
    mapred.reduce.tasks = 12
  ),
  copyFiles = TRUE,
  readback  = FALSE
)
```

In this execution function, we added two more things to `rhwatch` function. First we add `load` function to
load `label.RData` to the R global environment of each reducer. And we put this loading step in the `setup`
is because we would like to make `xlab` and `ylab` objects available in the reduce step. Second we add a new
argument `shared`, which is a character vector of files located on the HDFS. At the beginning of the MapReduce 
job, these files will be copied to the local hard disks of the mapper and reducer. This is a necessary step
for loading the `label.RData` right before the reduce step on reducer.

The difference between using `parameters` and `shared` argument to make specific objects available for every 
mapper and reducer is that how large shared objects are. If they are relative small size of objects like two
strings or a small vector, then `parameters` argument is prefered. On the other hand, if shared objects are
some large size of objects like a data.frame or matrix, then `shared` argument should be considered becasue
we put these shared objects on HDFS first and then grab them from HDFS to our map or reduce function.

```{r eval=FALSE, tidy=FALSE}
rhls("/ln/tongx/housing/listplotbystate/_outputs")
```
```
   permission owner      group     size          modtime                                                           file
1  -rw-r--r-- tongx supergroup 197.6 kb 2014-10-02 22:23 /ln/tongx/housing/listplotbystate/_outputs/list.vs.time.AL.pdf
2  -rw-r--r-- tongx supergroup 213.2 kb 2014-10-02 22:23 /ln/tongx/housing/listplotbystate/_outputs/list.vs.time.AR.pdf
3  -rw-r--r-- tongx supergroup 51.46 kb 2014-10-02 22:23 /ln/tongx/housing/listplotbystate/_outputs/list.vs.time.AZ.pdf
4  -rw-r--r-- tongx supergroup 192.2 kb 2014-10-02 22:23 /ln/tongx/housing/listplotbystate/_outputs/list.vs.time.CA.pdf
......
```
Still, all results of plotting will be located in `_outputs` of output directory.

```{r eval=FALSE, tidy=FALSE}
rhget("/ln/tongx/housing/listplotbystate/_outputs/list.vs.time.VA.pdf", "./")
rhget("/ln/tongx/housing/listplotbystate/_outputs/list.vs.time.CA.pdf", "./")
```
