### Manage RHIPE Jobs ###

As you submit more Rhipe jobs, you'll need to check their status and manage them.  In this short
section, we'll introduce two new functions to do just that. 

First, let's submit a simple job:

```{r eval=FALSE, tidy=FALSE}
map6 <- expression(
  while(TRUE) {}
)
mr6 <- rhwatch(
  map      = map6,
  input    = rhfmt("/ln/tongx/housing/housing.txt", type = "text"),
  output   = rhfmt("/ln/tongx/housing/tmp", type = "sequence"),
  mapred   = list(
    mapred.reduce.tasks = 0
  ),
  readback = FALSE,
  noeval   = TRUE
)
badjob <- rhex(mr6, async = TRUE) 
```

We've set `mapred.reduce.tasks` to be 0, because this job omits the optional reduce expression.  
The map contains an infinite loop, so this job will never complete. The only purpose for this meaningless
map expression is we want the job is keeping running when we try to access its status. But since we've run
this job on the backgroup, we can check its status with `rhstatus`, which takes as an argument
the object returned by `rhex`. When we said background, it means that the job is running on the 
Hadoop, and we still have the control ability of R session at the same time. 

```{r eval=FALSE, tidy=FALSE}
rhstatus(badjob)
```
```
[Tue Sep 30 00:34:20 2014] Name:2014-09-30 00:34:05 Job: job_201405301308_4753  State: RUNNING Duration: 14.673
URL: http://deneb.stat.purdue.edu:50030/jobdetails.jsp?jobid=job_201405301308_4753
             pct numtasks pending running complete killed failed_attempts
map    0.0191228        1       0       1        0      0               0
reduce 0.0000000        0       0       0        0      0               0
       killed_attempts
map                  0
reduce               0
Waiting 5 seconds
```

Type `ctrl+c` to stop the status updates.  Since we know this job won't ever complete, 
we'll have to kill it.  Again, use the object returned by `rhex`

```{r eval=FALSE, tidy=FALSE}
rhkill(badjob)
```

You can call `rhstatus` again to see that the job is now in state `KILLED`
