## The Cluster Performance Test ##

### Introduction ###

We'll demonstrate the computation ability of `RHIPE` package with a performance test by using logistic
regression. 

The basic idea is firstly, we wish to generate an observation sample with a size of *N*, *V* number 
of variables, which means there are *V-1* explanatory variables.  In summary, we wish to generate a 
dataframe with *N* rows and *V* collumns. Secondly, we want to use the logistic regression method to
analyze the dataframe. However, when the data size is quite large, it will be impractical to generate
and analyse the whole data set in terms of the time cosuming and the memory limitation.

Instead, we will use `RHIPE` package to generate subsets first and then use logistic regression 
method to analyze each subset by R function `glm.fit`. The subset observation sample size is *M* and
the number of subsets is *R*. As a result, $N = M * R$. Each subset contains a sub-dataframe with *M*
rows and *V* collumns. We will use R function `system.time` to measure elapsed time which will be 
explained more in the later session.

In this example, we will show how to do a sing run of the performance test, which means we only pick
one possible value of *M* and one possible value of *V* to see how fast it is to go. 

We choose $n = log2(N) = 25$, $v = log2(V) = 5$, $m = log2(M) = 13$, and $log2(R) = 12$ to illustrate
one single run. So the size for the whole dataset is $2^25 * 2^5 * 2^3 = 8GB$. We will generate each
subset with the size of $2^13 * 2^5 * 2^3 = 2MB$

### Generate Datasets ###

As we mentioned in the `Housing Data` example, the first step in a D/&R analysis is to choose a 
division method and create subsets. In this example, we devide the whole dataset to *R* sets, where
$ log2(R) = 12$. So we will generate 2^12 matrices with 2^13 rows and 2^5 columns each. 

Before we run the map function, we will set up some basic parameters in the front end of R enviroment:

```{r,eval=FALSE,tidy=FALSE}
n   <- 25
m   <- 13
v   <- 5
p   <- 2^v - 1
rep <- 1

```

Here we define an R object called `p`, the value of which equals to $2^v - 1$. As introduced in the
last part, $2^v$ is the number of variables, so `p` is the number of the explanatory variables. We 
introduce `p` here to make the following codes much easier to write and understand. We also specify
the replicate value `rep` is 1 because we only have a single run of the cluster performance test. But
in your future performance test, you might want to have more replicates for the test, and you can
easily change the value of `rep`.

#### Map ####

```{r, eval=FALSE,tidy=FALSE}
map1 <- expression({
  for (r in map.values){
    set.seed(r)
    value <- matrix(c(rnorm(m*p), sample(c(0,1), m, replace=TRUE)), ncol=p+1)
    rhcollect(r, value)
  }
})
```
 
In this case, we don't have an existing data which is different from the `Housing Data` example. 
Creating a dataset by randomly simulation is the part of the `map` job. The input keys and values are
the same by default for the sumulating data step. The keys are numeric numbers 1, 2, ...., *R*,  where
*R* is the number of the subsets. In our example, log2(R) = log2(N/M) = 12. They are the elements of
the list object `map.keys`, which in this case, `map.keys` and `map.values` would be the same. The 
output keys from the `map1` function are the same as the input keys. This is defined by the `map1` 
function. The output value for each key is a matrix with 2^13 rows and 2^5 columns. As we have
explained in the `Housing Data` example, the output key-value pairs from the `map` function 
are also called intermediate key-value pairs. So `rhcollect` emits these intermediate key-value 
pairs.

There is no reducer step for the generation part. Next we will write the generated subsets to HDFS.

#### Execution Function ####

```{r eval=FALSE, tidy=FALSE}
dir.dm  = "/ln/song273/tmp/multi.factor_n25/dm/n25v5m13"
mr1 <- rhwatch(
 map      = map1,
 input    = c(2^(n-m),12),
 output   = dir.dm,
 jobname  = dir.dm,
 mapred   = list(
   mapred.task.timeout=0,
   mapred.reduce.tasks=0),
 parameters = list(m = 2^m, p = p),
 readback = FALSE
  )
```

We have introduced the `rhwatch` function before. In this case, `rhwatch` function will submit our 
MapReduce jobs to finish the data generation part and write the subsets to HDFS. There are three types
of the `input` as Text file input, Sequence file input and lapply input. The first two types of input
have been introduced in the `Housing Data` example. And in our performance example, the type of 
`input` is lapply input because we are generating numeric data. The argument `input` in our 
performance test example is a vector consisting of two elements. The first element specifies the 
input key would be from 1 to $2^(n-m)$ for `map1` function. Also, it specifies how many tasks you 
would like Hadoop MapReduce Framework to run because each task conresponds to one key-value pair. 
In our performance test example, each task is to generate a subset with 2^13 rows and 2^5 columns. 
The second element of the vector `input` specifies the numbers of mappers you would like to use. So 
each mapper would run $2^(n-m)/12$ tasks, where `12` is the value of the second element of `input`. 
So the value of the second element would definitely affect the performance test and it is worthy to 
carry out an experiment to test its effect. 

Once you try the `Housing Data` example, you have run into most arguments of `rhwatch`. There are some
arguments of `rhwatch` are not been used at all which means they don't have a default value such as 
the argument `parameters`.  But in some cases, they are very important. The argument `parameters` is
a list with parameters to be passed to a mapreduce job. It is an very important argument because it 
passes all the possible values we will need in the Hadoop MapReduce jobs to HDFS. in our case, we 
specify the values of `m` and `p` in the front end R, but Hadoop doesn't have these values in the 
back end. So the argument `parameter` will packages `m` and `p` as a list and distribute it to HDFS.
The argument `jobname` is not neccesary to be set. It is the name of the job, which is visible on 
the Jobtracker website. If not provided, Hadoop MapReduce uses the default name job_date_time_number
e.g. job_201007281701_0274.

```
[Sun Oct  5 22:30:58 2014] Name:/ln/song273/tmp/multi.factor_n25/dm/n25v5m13 Job: job_201405301308_5050  State: RUNNING Duration: 30.455
URL: http://deneb.stat.purdue.edu:50030/jobdetails.jsp?jobid=job_201405301308_5050
             pct numtasks pending running complete killed failed_attempts
map    0.9999999       12       0      12        0      0               0
reduce 0.0000000        0       0       0        0      0               0
       killed_attempts
map                  0
reduce               0
Waiting 5 seconds

```

Our subsets have been created and are saved on HDFS now. We will use `rhls()` to see more details 
about files on HDFS.

```{r eval=FALSE, tidy=FALSE}
rhls("/ln/song273/tmp/multi.factor_n25/dm/n25v5m13")
```

```
   permission   owner      group     size          modtime
1  -rw-r--r-- song273 supergroup        0 2014-10-05 22:32     /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/_SUCCESS
2  drwxrwxrwt song273 supergroup        0 2014-10-05 22:30        /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/_logs
3  -rw-r--r-- song273 supergroup   690 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00000
4  -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00001
5  -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00002
6  -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00003
7  -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00004
8  -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00005
9  -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00006
10 -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00007
11 -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00008
12 -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00009
13 -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00010
14 -rw-r--r-- song273 supergroup   682 mb 2014-10-05 22:30 /ln/song273/tmp/multi.factor_n25/dm/n25v5m13/part-m-00011

```
It generates 12 files with ending name "00000" to "00011". Most of their size is 682 Mb. Each file
consists of $2^12 / 12$ subsets. 

### Elapsed Time Measurement ###

There are two types of elapsed-time computation. The subsets are stored on the HDFS as R objects. 
The first computation type is **O**, the elapsed time to read the subsets from the HDFS and make 
them available to `glm.fit` in memory as an R objects. The other type, **L**, starts when **O** ends
and it consists of `glm.fit` computations on the subsets by **map**, plus **reduce** gathering the 
subset estimates and computing the means. However, we cannot measure **L** directly. So we measure 
**O** in one run and **T = O + L** in another.

#### The First Kind Of Elapsed Time -- **O** ####

```{r eval=FALSE, tidy=FALSE}
timing   <- data.frame()
type  <- "O"
```

First, we will initialize a data frame called `timing` to store these two different kinds of elapsed time later and create a character `type` to lable **O** and **T** seperately. 
##### Map #####

```{r eval=FALSE, tidy=FALSE}
map2 <- expression({})
```

It is important to notice that `map2` is an expression with no command. That's because we only want
to read the data from HDFS to the front end without other operations and then record the reading time .

##### Execution Function #####

```{r eval=FALSE, tidy=FALSE}
dir.dm  <- "/ln/song273/tmp/multi.factor_n25/dm/n25v5m13"
dir.nf  <- "/ln/song273/tmp/multi.factor_n25/nf/n25v5m13"
mr2 <- rhwatch(
 map      = map2,
 input    = dir.dm,
 output   = dir.nf,
 jobname  = dir.dm,
 mapred   = list(
   mapred.reduce.tasks=0,
   rhipe_map_buff_size=2^15),
 parameters = list(p = p),
 noeval   = TRUE,
 readback = FALSE
  )
t      <- as.numeric(system.time({rhex(mr2, async=FALSE)})[3])
t      <- data.frame(type=type,n=n,m=m,v=v,rep=rep,t=t)
timing <- rbind(timing,t)
```
```
Saving 1 parameter to /tmp/rhipe-temp-params-c677abd27e3b1654c25dccadaa7e3483 (use rhclean to delete all temp files)

```
The `noeval` is another very useful argument for `rhwatch`. By default it is set to be 'FALSE'.
In our example, we set `noeval` equals to 'TRUE', which means the  Hadoop MapReduce job will not run,
and `rhwatch` just return an R object that contains all the information required by Rhipe to run a 
MapReduce job. Instead, the `rhex` function will submit the MapReduce job to the Hadoop MapReduce 
framework. 

The `dir.dm` and `dir.nf` specify the `input` and `output` argument of `rhwatch` function. 
So the first `t` in the R codes records the first type elapsed time **O** and the second `t` is a 
data frame which saves the value of computation `type`, the value of `n`, `m`, `v`, the number of 
replicate `rep`, and the value of first type of elapsed time **O**. Then we use R function `rbind` 
to combine data frame `timing` and data frame `t` by rows. We can take a look at the data frame 
`timing`.

```{r eval=FALSE, tidy=FALSE}
timing
```

```
```
#### The Second Kind of Elapsed Time -- **T** ####

```{r eval=FALSE, tidy=FALSE}
type  <- "T"
```

We first specify the value of `type` as `T` to identify the second kind of elapsed time **T**.

##### Map #####

```{r eval=FALSE, tidy=FALSE}
map3 <- expression({
 for (v in map.values) {
  value = glm.fit(v[,1:p],v[,p+1],family=binomial())$coef
  rhcollect(1, value)
 }
})
```

The input to our `map3` is the final output we created in the previous section which the key is from
1 to $2^13$ and the input value is a dataframe with $2^13$ rows and $2^5$ collumns. For each subset,
we apply the same action : take the coefficient of the logistic regression for the subset. They are 
stored in a single row dataframe to be recombined in the reduce step below. This single row dataframe
has `p` collumns. We use the same key for all the subsets so that they wil appear together in the 
reduce step. As before, `rhcollect` emits intermediate key-value pairs.

##### Reduce #####

```{r eval=FALSE, tidy=FALSE}

reduce3 <- expression(
  pre = {
    v = rep(0,p) 
    nsub = 0
  },
  reduce = {
    v = v + colSums(matrix(unlist(reduce.values), ncol=p, byrow=TRUE)) 
    nsub = nsub + length(reduce.values)
  },
  post = {
   rhcollect(reduce.key, v/nsub)
  }
)
```

In the reduce step, we recombine the coefficients from different subset together and sum them up by 
collumn using R function `colSums` and and get the estimate of the coefficients by the sample mean.

The input key is the placeholder value 1, which is left unchanged as the output key. The input values
are the single row data frames for each subset, and the output values is the single row data frames
consisting of the mean estimate for the coefficients. In `post`, we use `rhcollect` to emit the final
output so that it will be written to HDFS.

##### Exectution Function #####

```{r eval=FALSE, tidy=FALSE}
dir.dm  = "/ln/song273/tmp/multi.factor_n25/dm/n25v5m13"
dir.gf  = "/ln/song273/tmp/multi.factor_n25/gf/n25v5m13"

mr3 <- rhwatch(
  map      = map3,
  reduce   = reduce3,
  input    = dir.dm,
  output   = dir.gf,
  mapred   = list(
    mapred.reduce.tasks=1,
    rhipe_map_buff_size=10
  ),
 parameters = list(p=p),
 jobname    = dir.gf,
 noeval     = TRUE
)

t <- as.numeric(system.time({rhex(mr3, async=FALSE)})[3])
t = data.frame(type=type,n=n,m=m,v=v,rep=rep,t=t)
timing = rbind(timing,t)
```
```
Saving 2 parameters to /tmp/rhipe-temp-params-4c3905d5e34c9689c92490e17cb6162b (use rhclean to delete all temp files)
```
After the job completes successfully, we record the whole elapsed time **T** and save it to the R 
object `t`. And we save the results to our data frame `timing`. Till now, we have finished the single
run of the cluster performance test and save the result to a data frame called `timing`.

### Performance Test Results ###
```{r eval=FALSE, tidy=FALSE}
timing
```

```
     type  n    m    v    rep       t
1       O  23   13   5    1      55.120
2       T  23   13   5    1     147.923

```

As we can see from the `timing` dataframe, the elapsed time **O** we need to read the datasets 
from HDFS to the front end is 55.12 seconds and the whole elapsed time **T** which consists of the 
reading time and the analyzing time is 147.923 seconds, which is less than 2.5 mins. That result is 
impressive! The cluster we use here to run the test is a very small one with two nodes and 64 GB 
memory in total. You can definitely use RHIPE computation environment to generalize this one sing 
run to a full cluster performance test. If you want to explore more about the performance test, you 
can check the reference paper for more details. 



