# RHIPE Tutorial #

## Installation ##

### Installation ###
RHIPE is an R package, that can be downloaded at [this website](http://ml.stat.purdue.edu/rhipebin/Rhipe_0.73.1.tar.gz). 
To install it the user needs to
- Set an environment variable *HADOOP* that points to the Hadoop installation directory. It is 
expected that *HADOOP\bin* contains the Hadoop shell executable *hadoop*.
- A version of Google's Protocol Buffers [here](https://code.google.com/p/protobuf/) greater than
2.3.0

Once the package has been downloaded the user can install it via

```{r,eval=FALSE}
R CMD INSTALL Rhipe_version.tar.gz
```

where *version* is the latest version of RHIPE. The source is under version control at [GitHub](https://github.com/tesseradata/RHIPE/).

This needs to be installed on all the computers: the one you run your R environment and all the task
computers. Use RHIPE is much easier if your filesystem layout (i.e location of R, Hadoop, libraries 
etc) is identical across all computers.

### Tests ###

In R

```{r,eval=FALSE}
library(Rhipe)
rhinit()
```

should work successfully.

```{r,eval=FALSE}
## a is a list consisting of 3 elements and each element is a list containing two elements
a = list(list(1,c(1,2,3)),list(2,c(1,2,3)),list(3,c(1,2,3)))
rhwrite(a,"/tmp/xx")
```

should successfully write the list to the HDFS

```{r,eval=FALSE}
b = rhread("/tmp/xx")
str(b)
```

Return a list of length 3 each element a list of 2 objects.

```
List of 3
 $ :List of 2
  ..$ : num 1
  ..$ : num [1:3] 1 2 3
 $ :List of 2
  ..$ : num 2
  ..$ : num [1:3] 1 2 3
 $ :List of 2
  ..$ : num 3
  ..$ : num [1:3] 1 2 3
```

And a quick run of this should also work

```{r,eval=FALSE}
map <- expression({
  lapply(seq_along(map.values),function(r){
    x <- runif(map.values[[r]])
    rhcollect(map.keys[[r]],c(n = map.values[[r]],mean = mean(x),sd = sd(x)))
  })
})

## Create a job object
z <- rhwatch(
  map    = map, 
  input  = c (10,10),
  output = "/tmp/xx", 
  mapred = list(mapred.reduce.tasks=0),
  noeval = TRUE
  )

## Submit the job
rhex(z, async = FALSE )
## Read the results
res    <- rhread('/tmp/xx')
colres <- do.call('rbind',lapply(res,"[[",2))
colres
```

```
       n      mean        sd
 [1,]  1 0.4314606        NA
 [2,]  2 0.1920560 0.2477057
 [3,]  3 0.5399694 0.3059996
 [4,]  4 0.5103296 0.3543939
 [5,]  5 0.6566606 0.2697990
 [6,]  6 0.5678072 0.2698144
 [7,]  7 0.5696185 0.2967622
 [8,]  8 0.5207922 0.3553869
 [9,]  9 0.5791976 0.3101113
[10,] 10 0.4795422 0.2394009


```

