### Hadoop as a Queryable Database ###

*Sightly artificial:* store all American Airlines information indexed by year, month, and day.
Each (year, month, day) triplet will have all flight entries that left on that day. Using the above
data set as the source, the Southwest lines are selected and sent to the reducer with the (year, 
month,day) key. All flights with the same (year, month) will belong to the same file. Given a (year
, month,day) triplet, we can use the Map File output format to access the associated flight 
information in seconds rather than subsetting using Map-reduce.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  h <- do.call("rbind", map.values)
  d <- h[h$carrier == 'AA', , drop = FALSE]
  if(nrow(d) > 0) {
    e <- split(d, list(d$year, d$month, d$day))
    lapply(e, function(r) {
      k <- as.vector(unlist(r[1, c("year", "month", "day")]))  ## remove attributes
      rhcollect(k, r)
    })
  }
})
reduce <- expression(
  pre = { 
    collec <- NULL 
  },
  reduce = {
    collec <- rbind(collec, do.call("rbind", reduce.values))
    collec <- collec[order(collec$depart), ]
  },
  post = {
    rhcollect(reduce.key, collec)
  }
)
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks", type = "sequence"),
  output   = rhfmt("/tmp/airline/output/southwest", type = "map"),
  mapred   = list(rhipe_map_buff_size = 10),
  readback = FALSE
)
```

Attributes are removed in line $8$, for otherwise we have to retrieve a data frame with a data frame 
with column names and row names instead of a more convenient numeric vector. The map expression 
combines the individual data frames. Each data frame has $5000$ rows, hence `rhipe_map_buff_size` is 
set to 10 for a combined data frame of $50000$ rows in line $32$. This is crucial. The default value 
for `rhipe_map_buff_size` is $10,000$. Binding $10,000$ data frames of $5000$ rows each creates a data 
frame of 50MN rows - too unwieldy to compute with in R (for many types of operations). Data frames 
for American Airlines (*carried code = AA*) are created and emitted with the call to rhcollect in line
$15$. These are combined in the reduce since data frames for the same (year, month, day) triplet can 
be emitted from different map expressions. Since this is associative and commutative we use a 
combiner. The output format (`inout[[2]]`) is map, so we can access the flights for any triplet with 
a call to `rhmapfile` which returns a list of key,value lists.

```{r eval=FALSE, tidy=FALSE}
x <- rhmapfile("/tmp/airline/output/southwest")
x[[c(1987,10,2)]]
```