<!--
### Streaming Data ###

Some algorithms are left associative in their operands $t_1$, $t_2$, $\ldots$, $t_n$ but not 
commutative. For example a streaming update algorithm that computes the inter-arrival times 
of time series data for different levels of a categorical variable. That is, the triangular 
series $t_{k,1}$, $t_{k,2}$, $\ldots$, $t_{k,n_k}$ where k takes the levels of a categorical 
variable C (which takes the values $1, 2, 3, \ldots, m$). The input are pairs (i, j), $i$ $\in$ 
$\{1, 2, \ldots, m\}$, $j$ $\in$ $\{t_{ik}\}$. In the following code, the data structure F 
is updated with the datastructure contained in the values (see the map). The datastructures
are indexed in time by the timepoint - they need to be sent to the reducer (for a given level
of the categorical variable catlevel) in time order. Thus the map sends the pair (catlevel, 
timepoint) as the key. By using the part parameter all the data structures associated with the 
catlevel are sent to the same R reduce process. This is vital since all the component R expressions
in the reduce are run in the same process and namespace. To preserve numeric ordering we insist on
the special map output key class. With this special key class, we cannot have a map output format. 
In the reduce, the setup expression redsetup is run upon R startup (the process assigned to several 
keys and their associated values). Then for each new intermediate key (catlevel, timepoint), it runs
the pre, reduce and post. The lack of a post is because we have exactly one intermediate value for 
a given key (assuming the time points for a category are unique). The redclose expression is run 
when all keys and values have been processed by the reducer and R is about to quit.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  lapply(seq_along(map.values),function(r){
    catlevel <- map.keys[[r]] #numeric
    timepoint <- map.values[[r]]$timepoint #numeric
    datastructure <- map.values[[r]]$data
    key <- c(catlevel, timepoint)
    rhcollect(key, datastructure)
  })
})
redsetup <- expression({
  currentkey <- NULL
 })
reduce <- expression(
  pre = {
    catlevel <- reduce.key[1]
    time <- reduce.key[2]
    if(!identical(catlevel,currentkey)){
    ## new categorical level
    ## so finalize the computation for
    ## the previous level e.g. use rhcollect
    if(!identical(currentkey, NULL))
      FINALIZE(F)
      ## store current categorical level
      currentkey <- catlevel
      ## initialize computation for new level
      INITIALIZE(F)
    }
  },
  reduce = {
    F <- UPDATE(F, reduce.values[[1]])
  }
)
redclose <- expression({
  ## need to run this, otherwise the last catlevel
  ## will not get finalized
  FINALIZE(F)
})
mr <- rhwatch(..., 
  combiner    = FALSE,
  setup       = list(reduce = redsetup),
  cleanup     = list(reduce = redclose),
  orderby     = "numeric",
  partitioner = list(lims = 1, type = "numeric")
)
```

#### Concrete (but artifical) Example ####

We will create a data set with three columns: the level of a categorical variable A, a time variable
B and a value C. For each level of A, we want the sum of differences of C ordered by B within A.

Creating the Data set The column A is the key, but this is not important. There are 5000 levels of A,
each level has 10,000 observations. By design the values of B are randomly written (sample), also for
simplicity C is equal to B, though this need not be.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  N <- 10000
  for(first.col in map.values) {
    w <- sample(N, N, replace = FALSE)
    for(i in w) {
      rhcollect(first.col, c(i, i))
    }
  }
})
mapred <- list(mapred.reduce.tasks = 0)
mr <- rhwatch(
  map = map, 
  input = c(5000, 10),
  output = rhfmt("/tmp/airline/output/sort", type = "sequence"),
  mapred = mapred,
  readback = FALSE,
  noeval = TRUE
)
rhex(mr)
```

Sum of Differences The key is the value of A and B, the value is C.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  lapply(seq_along(map.values),function(r){
    f <- map.values[[r]]
    rhcollect(as.integer(c(map.keys[[r]], f[1])), f[2])
  })
})
```

Thus each output from a map is a key (assuming there are not any duplicates for B for a given level
of A), thus reduce.values has only one observation. All keys sharing the same level of A will be 
sent to one R process and the tuples as.integer(c(map.keys[[r]],f[1])) will be sorted. reduce.setup 
is called once when the R process starts processing its assigned partition of keys and reduce.post 
is called at the end (when all keys have been processed)

```{r eval=FALSE, tidy=FALSE}
reduce.setup <- expression({
  newp <- -Inf
  diffsum <- NULL
})
reduce <- expression(
  pre = {
    if(reduce.key[[1]][1] != newp) {
      if(newp > -Inf) rhcollect(newp, diffsum) #prevents -Inf from being collected
      diffsum <- 0
      lastval <- 0
      newp <- reduce.key[[1]][1]
      skip <- TRUE
    }
  },
  reduce = {
    current <- unlist(reduce.values) #only one value!
    if(!skip) {
      diffsum <- diffsum + (current-lastval) 
    }else {
      skip <- FALSE
    }
    lastval <- current
  }
)
reduce.post <- expression({
  if(newp > -Inf) rhcollect(newp, diffsum) #for the last key
})
mr <- rhwatch(
  map = map,
  reduce = reduce, 
  input = rhfmt("/tmp/airline/output/sort", type = "sequence"),
  output = rhrmt("/tmp/airline/output/sort2", type =  "sequence"),
  partitioner = list(lims = 1, type = "integer"),
  orderby = "integer",
  cleanup = list(reduce = reduce.post),
  setup = list(reduce = reduce.setup),
  readback = FALSE
)
```

-->
### Simple Debugging ###

Consider the example code used to compute the delay quantiles by month (see Delay Quantiles By 
Month ). We can use tryCatch for some simple debugging. See the error in line 7, there is no such v
ariable isdelayed

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  tryCatch(
    {
      a <- do.call("rbind",map.values)
      a$delay.sec <- as.vector(a[,'arrive'])-as.vector(a[,'sarrive'])
      a <- a[!is.na(a$delay.sec),]
      a$isdelayed <- sapply(a$delay.sec,function(r) if(r>=900) TRUE else FALSE)
      a <- a[isdelayed==TRUE,] ## only look at delays greater than 15 minutes
      apply(a[,c('month','delay.sec')],1,function(r){
        k <- as.vector(unlist(r))
        if(!is.na(k[1])) rhcollect(k,1) # ignore cases where month is missing
      })
    },
    error = function(e) {
      e$message <- sprintf("Input File:%s\nAttempt ID:%s\nR INFO:%s",
        Sys.getenv("mapred.input.file"),
        Sys.getenv("mapred.task.id"),
        e$message
      )
      stop(e) ## WONT STOP OTHERWISE
    }
  )
})
reduce <- expression(
  pre = {
    sums <- 0
  } ,
  reduce = {
    sums <- sums + sum(unlist(reduce.values))
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/", type = "sequence"),
  output   = rhfmt("/tmp/airline/output/quantiledelay", type = "sequence"),
  mapred   = mapred,
  readback = FALSE
)
```

Produces a slew of errors like (output slightly modified to fit page)

```
03/07/14 00:41:20 INFO mapred.JobClient: Task Id :
  attempt_201007281701_0273_m_000023_0, Status : FAILED
java.io.IOException: MROutput/MRErrThread failed:java.lang.RuntimeException:
R ERROR
=======
Error in `[.data.frame`(a, isdelayed == TRUE, ) : Input File:
Attempt ID:attempt_201007281701_0273_m_000023_0
R INFO:object "isdelayed" not found
```

It can be very useful to provide such debugging messages since R itself doesn’t provide much help. 
Use this to provide context about variables, such printing the first few rows of relevant data 
frames (if required). Moreover, some errors don’t come to the screen instead the job finishes 
successfully (but very quickly since the R code is failing) but the error message is returned as a 
counter. The splits succeed since Hadoop has finished sending its data to R and not listening to 
for errors from the R code. Hence any errors sent from R do not trigger a failure condition in 
Hadoop. This is a RHIPE design flaw. To compensate for this, the errors are stored in the counter 
`R_ERROR`.

Unfortunately, RHIPE does not offer much in the way of debugging. To run jobs locally that is, 
Hadoop will execute the job in a single thread on one computer, set `mapred.job.tracker` to local in 
the `mapred` argument of `rhwatch`. In this case, `shared` cannot be used and `copyFiles` will 
not work.

