### Analysis ###

We compute some summaries and displays to understand the data.

#### Top 20 cities by total volume of flights ####

We compute some summaries and displays to understand the data.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind", map.values)
  inbound <- table(a[, 'origin'])
  outbound <- table(a[, 'dest'])
  total <- table(unlist(c(a[, 'origin'], a['dest'])))
  for(n in names(total)) {
    inb <- if(is.na(inbound[n])) 0 else inbound[n]
    ob <- if(is.na(outbound[n])) 0 else outbound[n]
    rhcollect(n, c(inb, ob, total[n]))
  }
})
reduce <- expression(
  pre = {
    sums <- c(0, 0, 0)
  },
  reduce = {
    sums <- sums + apply(do.call("rbind", reduce.values), 2, sum)
  },
  post = {
    rhcollect(reduce.key, sums)
  }
)
mapred <- list(rhipe_map_buff_size = 15)
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks", type = "sequence"),
  output   = rhfmt("/tmp/airline/output/volume", type = "sequence"),
  mapred   = mapred,
  readback = FALSE
)
```

The code is straightforward. We increased the value of `rhipe_map_buff_size` since we are doing 
summaries of columns. The figure 
[Log of time to complete vs log of rhipe_map_buff_size](./plots/rhipemapbuff.pdf). Plots the
time of completion vs the mean of three trials for different values of `rhipe_map_buff_size`. The 
trials set `rhipe_map_buff_size` to 5, 10, 15, 20, 25 and 125. All experiments (like the rest in the 
manual) were performed on a 72 core cluster across 8 servers with RAM varying from 16 to 64 GB.

Read the data into R and display them using the lattice library.

```{r eval=FALSE, tidy=FALSE}
counts <- rhread("/airline/volume")
aircode <- unlist(lapply(counts, "[[",1))
count <- do.call("rbind",lapply(counts,"[[",2))
results <- data.frame(
           aircode = aircode,
               inb = count[, 1],
               oub = count[, 2],
               all = count[, 3],
  stringsAsFactors = FALSE
)
results <- results[order(results$all, decreasing = TRUE), ]
library(lattice)
r <- results[1:20, ]
af <- reorder(r$aircode, r$all)
trellis.device(postscript, file = "volume.ps", color=TRUE, paper="letter")
dotplot(af ~ log(r[, 'all'], 10),
  xlab = 'Log_10 Total Volume',
  ylab = 'Airport',
   col = 'black',
  aspect = 1
)
dev.off()
```

There are 352 locations (airports) of which the top 20 serve 50% of the volume.
[Dot plot of top 20 cities by total volume of flights](./plots/volume.pdf)

#### Carrier Popularity ####

Some carriers come and go, others demonstrate regular growth. In the following display, the log base
10 volume (total flights) over years are displayed by carrier. The carriers are ranked by their 
median volume (over the 10 year span).

As mentioned before, RHIPE is mostly boilerplate. Notice the similarities between this and previous
examples (on a side note, to do this for 12GB of data takes 1 minute and 32 seconds across 72 cores 
and all the examples, except the download and conversion to R data frames, in the manual are less 
than 10 minutes)

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind", map.values)
  total <- table(years = a[, 'year'], a[, 'carrier'])
  ac <- rownames(total)
  ys <- colnames(total)
  for(yer in ac){
    for(ca in ys){
      if(total[yer, ca] > 0) 
        rhcollect(c(yer, ca), total[yer, ca])
    }
  }
})
reduce <- expression(
  pre = {
    sums <- 0
  },
  reduce = {
    sums <- sums + sum(do.call("rbind", reduce.values))
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)

mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/", type="sequence"),
  output   = rhfmt("/tmp/airline/output/carrier.pop/", type="sequence"),
  mapred   = mapred,
  readback = FALSE
)
```

This is the RHIPE code to create summaries. We need to extract the data from Hadoop and create a 
display.
[Carrier Popularity](./plots/carrier.pdf)

```{r eval=FALSE, tidy=FALSE}
a <- rhread("/tmp/airline/output/carrier.pop")
head(a, 3)
```
```
[[1]]
[[1]][[1]]
[1] "1987" "AA"  

[[1]][[2]]
[1] 165121


[[2]]
[[2]][[1]]
[1] "1987" "AS"  

[[2]][[2]]
[1] 21406


[[3]]
[[3]][[1]]
[1] "1987" "CO"  

[[3]][[2]]
[1] 123002
```

```{r eval=FALSE, tidy=FALSE}
yr <- as.numeric(unlist(lapply(lapply(a, "[[", 1), "[[", 1)))
carrier <- unlist(lapply(lapply(a, "[[", 1), "[[", 2))
count <- unlist(lapply(a, "[[", 2))
results <- data.frame(
                yr = yr,
           carcode = carrier,
             count = count,
  stringsAsFactors = FALSE
)
results <- results[order(results$yr, results$count, decreasing = TRUE), ]
carr <- reorder(results$carcode, results$count, median)
trellis.device(postscript, file = "carrier.ps", color=TRUE, paper="letter")
xyplot(log(count, 10) ~ yr | carr, 
  data           = results,
  xlab           = "Years", 
  ylab           = "Log10 count",
  col            = 'black',
  scales         = list(scale = 'free',tck = 0.5,cex = 0.7),
  layout         = c(4, 4),
  aspect         = " xy",
  type           = 'b',
  par.strip.text = list(lines = 0.8, cex = 0.7), 
  cex            = 0.5,
  panel          = function(...){
    panel.grid(h = -1,v = -1)
    panel.xyplot(...)
  }
)
dev.off()
```

#### Proportion of Flights Delayed ####

It is very likely in the future analysis, we want to study the flights information for a specific 
day.So for this scenario we want to create new key/value pairs by using RHIPE. The input files are
the blocks of data we created previously, and the ouput will be 'sequence' file with key is the 
date, and corresponding value is a data frame of data for that particular day. For example, we would
like to know what is the delay rate on everyday.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind", map.values)
  a$delay.sec <- as.vector(a[,'arrive']) - as.vector(a[,'sarrive'])
  a <- a[!is.na(a$delay.sec),]
  a$isdelayed <- sapply(a$delay.sec, function(r) if(r > 900) TRUE else FALSE)
  e <- split(a, list(a$year, a$month, a$day))
  lapply(e, function(r){
    n <- nrow(r)
    numdelayed <- sum(r$isdelayed)
    rhcollect(as.vector(unlist(c(r[1, c("year", "month", "day")]))), c(n, numdelayed))
  })
})
reduce <- expression(
  pre = {
    sums <- c(0, 0)
  },
  reduce = {
    sums <- sums + apply(do.call("rbind", reduce.values), 2, sum)
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks", type = "sequence"),
  output   = rhfmt("/tmp/airline/output/delaybyday", type = "sequence"),
  mapred   = mapred,
  readback = FALSE
)
```

In map expression, we still iterate over the `map.values`, which are number `rhipe_map_buff_size` 
of the data frame with 5000 rows and 12 columns flights information. we define the second of delay
as `delay.sec`. Of course, we have to remove all `NA` in `delay.sec` since there are some records 
of flight have `NA` as missing data of arriving time. Then create a flag variable `isdelayed` to 
identify if the flight is delayed. Object `e` is a list which come from the calling of `split()` 
function. What we get for `e` is a data frame for each day as elements of the list. At last, we
collect the key which is the date, and value which is a vector with total number of flights and 
number of delayed flights for each element of `e`.

In reduce expression, we initialize the `sums` in `pre` of reduce, which will be the final total
number of flights and number of delay for a given day. And in `reduce` of reduce, we just cumulate
all two numbers for same key. Finally, in `post` of reduce, collect the final key/value pairs. 
`reduce.key` here is one particular date of the day, and `reduce.values` is a list with all 
`c(numberofflight, numberofdelay)` as elements.

We read the output by using `rhread()`, and then grab all the keys assigned to `y1`, grab all the 
values assigned to `y2`. Based on keys and values, we create a data frame named `results` with 6 
columns. The delay rate is the number of delay divided by the number of total flights on that day. 
Finally, the data frame is sorted by the day.

```{r eval=FALSE, tidy=FALSE}
b <- rhread("/tmp/airline/output/delaybyday")
y1 <- do.call("rbind", lapply(b, "[[", 1))
y1 <- y1[-1, ]
y2 <- do.call("rbind", lapply(b, "[[", 2))
y2 <- y2[-1, ]
results <- data.frame(
  year    = y1[, 1],
  month   = y1[, 2],
  day     = y1[, 3],
  nflight = y2[, 1],
  ndelay  = y2[, 2]
)
results$rate <- results$ndelay/results$nflight
results <- results[order(results$year, results$month, results$day), ]
head(results)
```
```
   year month day nflight ndelay      rate
1  1987    10   1   14759   9067 0.6143370
10 1987    10  10   13417   7043 0.5249311
11 1987    10  11   14016   7790 0.5557934
12 1987    10  12   14792   8376 0.5662520
13 1987    10  13   14859   8623 0.5803217
14 1987    10  14   14799   8806 0.5950402
```

STL decomposition of proportion of flights delayed is the STL decomposition of p (the proportion of
flights delayed). The seasonal panel clearly demonstrates the holiday effect of delays. They don’t 
seem to be increasing with time (see trend panel).

```{r eval=FALSE, tidy=FALSE}
prop <- results[,'rate']
prop <- prop[!is.na(prop)]
tprop <- ts(log(prop/(1 - prop)), 
  start     = c(1987, 273),
  frequency = 365
)
tprop[is.infinite(tprop)] <- 0
trellis.device(postscript, file = "propdelayedxyplot.ps", color=TRUE, paper="letter")
plot(stl(tprop,s.window="periodic"))
dev.off()
```

[STL decomposition of proportion of flights delayed](./plots/propdelayedxyplot.pdf)

We can do very similar thing that calculating the delay rate, but for each hour, instead of doing 
that for each day. We only need to change the key to be the hour variable in data.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind",map.values)
  a$delay.sec <- as.vector(a[, 'arrive'])-as.vector(a[, 'sarrive'])
  a <- a[!is.na(a$delay.sec),]
  a$isdelayed <- sapply(a$delay.sec, function(r) if(r > 900) TRUE else FALSE)
  a$hrs <- as.numeric(format(a[, 'sdepart'], "%H"))
  e <- split(a,a$hrs)
  lapply(e, function(r){
    n <- nrow(r) 
    numdelayed <- sum(r$isdelayed)
    rhcollect(as.vector(unlist(c(r[1, c("hrs")]))), c(n, numdelayed))
  })
})
reduce <- expression(
  pre = {
    sums <- c(0, 0)
  },
  reduce = {
    sums <- sums + apply(do.call("rbind", reduce.values), 2, sum)
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/", type = "sequence"),
  output   = rhfmt("/tmp/airline/output/delaybyhours", type = "sequence"),
  mapred   = mapred,
  readback = FALSE
)
```

Does the delay proportion change with hour? It appears it does (see Proportion of flights delayed
by hour of day). The hours are scheduled departure times. Why are so many flights leaving in the 
hours (12-3) delayed?

```{r eval=FALSE, tidy=FALSE}
b <- rhread("/tmp/airline/output/delaybyhours")
y1 <- do.call("rbind", lapply(b, "[[", 1))
y2 <- do.call("rbind", lapply(b, "[[", 2))
results <- data.frame(
  hour = y1,
  rate = y2[, 2] / y2[, 1]
)
trellis.device(postscript, file = "delaybyhours.ps", color=TRUE, paper="letter")
dotplot( hour ~ rate, 
  data           = results,
  xlab           = "Proportion of Flights Delayed > 15 minutes", 
  ylab           = "Hour of Day",
  col            = 'black',
  aspect         = " xy"
)
dev.off()
```

[Proportion of flights delayed by hour of day](./plots/delaybyhours.pdf)

