#### Distribution of Delays ####

Summaries are not enough and for any sort of modeling we need to look at the distribution of the 
data. So onto the quantiles of the delays. We will look at delays greater than $15$ minutes. To 
compute approximate quantiles for the data, we simply discretize the delay and compute a frequency 
count for the unique values of delay. This is equivalent to binning the data. Given this frequency 
table we can compute the quantiles.

The distribution of the delay in minutes does not change significantly over months.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind", map.values)
  a$delay.sec <- as.vector(a[,'arrive']) - as.vector(a[,'sarrive'])
  a <- a[!is.na(a$delay.sec), ]
  a$isdelayed <- sapply(a$delay.sec, function(r) if(r >= 900) TRUE else FALSE)
  a <- a[a$isdelayed == TRUE, ] ## only look at delays greater than 15 minutes
  apply(a[, c('month', 'delay.sec')], 1, function(r) {
    k <- as.vector(unlist(r))
    if(!is.na(k[1])) rhcollect(k,1) # ignore cases where month is missing
  })
})
reduce <- expression(
  pre = {
    sums <- 0
  } ,
  reduce = {
    sums <- sums + sum(unlist(reduce.values))
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/",type="sequence"),
  output   = rhfmt("/tmp/airline/quantiledelay",type="sequence"),
  mapred   = mapred
  readback = FALSE
)
b <- rhread("/tmp/airline/quantiledelay")
y1 <- do.call("rbind",lapply(b, "[[", 1))
count <- do.call("rbind",lapply(b, "[[", 2))
results <- data.frame(
  month = y1[,1], 
  n     = y1[,2], 
  count = count
)
results <- results[order(results$month, results$n), ]
results.2 <- split(results, results$month)

discrete.quantile<-function(x, n, prob = seq(0,1,0.25), type = 7) {
  sum.n <- sum(n)
  cum.n <- cumsum(n)
  np <- if(type==7) (sum.n-1)*prob + 1 else sum.n*prob + 0.5
  np.fl <- floor(np)
  j1 <- pmax(np.fl, 1)
  j2 <- pmin(np.fl+1, sum.n)
  gamma <- np-np.fl
  id1 <- unlist(lapply(j1, function(r) seq_along(cum.n)[r <= cum.n][1]))
  id2 <- unlist(lapply(j2, function(r) seq_along(cum.n)[r <= cum.n][1]))
  x1 <- x[id1]
  x2 <- x[id2]
  qntl <- (1 - gamma)*x1 + gamma*x2
  qntl
}

DEL <- 0.05
results.3 <- lapply(seq_along(results.2), function(i) {
  r <- results.2[[i]]
  a <- discrete.quantile(r[, 2], r[, 3], prob = seq(0, 1, DEL))/60
  data.frame(
    month = as.numeric(rep(names(results.2)[[i]], length(a))),
    prop  = seq(0, 1, DEL),
    qt    = a
  )
})
results.3 <- do.call("rbind",results.3)
results.3$month <- factor(
  results.3$month,
  label = c("Jan","Feb","March","Apr","May","June",
    "July","August","September","October","November","December")
)
xyplot(log(qt,2) ~ prop | month, 
  data   = results.3,
  cex    = 0.4,
  col    = 'black',
  scales = list(x = list(tick.number = 10), y = list(tick.number = 10)),
  layout = c(4,3),
  type   = 'l',
  xlab   = "Proportion",
  ylab   = "log_2 delay (minutes)",
  panel  = function(x,y, ...){
    panel.grid(h = -1, v = -1)
    panel.xyplot(x, y, ...)
  }
)
```

[Quantiles by month](./plots/quantiles_by_month.pdf)

We can display the distribution by hour of day. The code is almost nearly the same. Differences are
in line $8$, where the hrs is used as the conditioning. But the results are more interesting. The 
delay amounts increase in the wee hours (look at panel $23,24,1,2$ and $3$)

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind",map.values)
  a$delay.sec <- as.vector(a[, 'arrive']) - as.vector(a[, 'sarrive'])
  a <- a[!is.na(a$delay.sec), ]
  a$isdelayed <- sapply(a$delay.sec, function(r) if(r >= 900) TRUE else FALSE)
  a <- a[a$isdelayed == TRUE,] ## only look at delays greater than 15 minutes
  a$hrs <- as.numeric(format(a[, 'sdepart'],"%H"))
  apply(a[, c('hrs','delay.sec')], 1, function(r) {
    k <- as.vector(unlist(r))
    if(!is.na(k[1])) rhcollect(k, 1)
  })
})
reduce <- expression(
  pre = {
    sums <- 0
  } ,
  reduce = {
    sums <- sums + sum(unlist(reduce.values))
  },  
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/",type="sequence"),
  output   = rhfmt("/tmp/airline/quantiledelaybyhour",type="sequence"),
  mapred   = mapred,
  readback = FALSE
)

```

The distribution of delay times by airports. This could be analyzed for several airports, but we 
take the top 3 in terms of volumes. In this display, the quantiles of log_2 of the delay times 
(in minutes) for inbound and outbound for 4 different airports is plotted. The airports are in 
order of median delay time. Of note, the median delay time for Chicago (ORD) and San Francisco 
(SFO) is greater flying in than out (approximately an hour). For both Chicago and Dallas Fort Worth
(DFW), the 75th percentile of inbound delays is greater than that for outbound. Quantile of minute 
delay for inbound and outbound for 4 different airports. Dotted red lines are 25%,50% and 75% 
uniform proportions. displays these differences.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  cc <- c("ORD","SEA","DFW","SFO")
  a <- do.call("rbind",map.values)
  a <- a[a$origin %in% cc| a$dest %in% cc,]
  if(nrow(a)>0){
    a$delay.sec <- as.vector(a[,'arrive'])-as.vector(a[,'sarrive'])
    a <- a[!is.na(a$delay.sec),]
    a$isdelayed <- sapply(a$delay.sec, function(r) if(r >= 900) TRUE else FALSE)
    a <- a[a$isdelayed == TRUE,]
    for(i in 1:nrow(a)){
      dl <- a[i, "delay.sec"]
      if(a[i,"origin"] %in% cc) {
        rhcollect(data.frame(dir = "outbound", ap = a[i,"origin"], delay = dl, stringsAsFactors = FALSE), 1)
      }
      if(a[i,"dest"] %in% cc) {
        rhcollect(data.frame(dir = "inbound",ap = a[i,"dest"], delay = dl, stringsAsFactors = FALSE), 1)
      }
    }
  }
})
reduce <- expression(
  pre = {
    sums <- 0
  } ,
  reduce = {
    sums <- sums + sum(unlist(reduce.values))
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/", type = "sequence"),
  output   = rhfmt("/tmp/airline/inoutboundelay", type = "sequence"),
  mapred   = mapred,
  readback = FALSE
)
```

#### Carrier Delays ####

Is there a difference in carrier delays? We display the time series of proportion of delayed 
flights by carrier, ranked by carrier.

```{r eval=FALSE, tidy=FALSE}
## For proportions and volumes
map <- expression({
  a <- do.call("rbind", map.values)
  a$delay.sec <- as.vector(a[,'arrive']) - as.vector(a[,'sarrive'])
  a <- a[!is.na(a$delay.sec),]
  a$isdelayed <- sapply(a$delay.sec,function(r) if(r >= 900) TRUE else FALSE)
  a$hrs <- as.numeric(format(a[,'sdepart'],"%H"))
  e <- split(a,a$hrs)
  lapply(e,function(r){
    n <- nrow(r) 
    numdelayed <- sum(r$isdelayed)
    rhcollect(as.vector(unlist(c(r[1,c("carrier")]))), c(n, numdelayed))
  })
})
reduce <- expression(
  pre = {
    sums <- c(0,0)
  },
  reduce = {
    sums <- sums + apply(do.call("rbind", reduce.values), 2, sum)
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
```

Proportion of flights delayed by carrier. Compare this with the previous graph.

#### Busy Routes ####

Which are busy the routes? A simple first approach (for display purposed) is to create a frequency 
table for the unordered pair (i,j) where i and j are distinct airport codes. Displays this over 
the US map.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  a <- do.call("rbind",map.values)
  y <- table(apply(a[,c("origin","dest")], 1, function(r){
    paste(sort(r),collapse=",")
  }))
  for(i in 1:length(y)){
    p <- strsplit(names(y)[[i]], ",")[[1]]
    rhcollect(p, y[[1]])
  }
})
reduce <- expression(
  pre = {
    sums <- 0
  },
  reduce = {
    sums <- sums + sum(unlist(reduce.values))
  },
  post = { 
    rhcollect(reduce.key, sums) 
  }
)
mapred <- list()
mapred$rhipe_map_buff_size <- 5
mapred$mapred.job.priority="VERY_LOW"
z <- rhwatch(
  map      = map,
  reduce   = reduce,
  combiner = TRUE,
  input    = rhfmt("/tmp/airline/output/blocks/", type = "sequence"),
  output   = rhfmt("/tmp/airline/ijjoin", type = "sequence"),
  mapred   = mapred,
  readback = FALSE
)

##Merge results
b <- rhread("/tmp/airline/ijjoin")
y <- do.call("rbind", lapply(b, "[[", 1))
results <- data.frame(
  a                = y[, 1],
  b                = y[, 2],
  count            = do.call("rbind", lapply(b, "[[", 2)),
  stringsAsFactors = FALSE
)
results <- results[order(results$count, decreasing=TRUE), ]
results$cumprop <- cumsum(results$count) / sum(results$count)
a.lat <- t(sapply(results$a, function(r) {
  ap[ap$iata == r, c('lat', 'long')]
}))
results$a.lat <- unlist(a.lat[, 'lat'])
results$a.long <- unlist(a.lat[, 'long'])
b.lat <- t(sapply(results$b,function(r) {
  ap[ap$iata == r,c('lat', 'long')]
}))
b.lat["CBM", ] <- c(0,0)
results$b.lat <- unlist(b.lat[,'lat'])
results$b.long <- unlist(b.lat[,'long'])
head(results)
```
```
   a   b  count     cumprop    a.lat     a.long    b.lat     b.long
1 ATL ORD 145810 0.001637867 33.64044  -84.42694 41.97960  -87.90446
2 LAS LAX 140722 0.003218581 36.08036 -115.15233 33.94254 -118.40807
3 DEN DFW 140258 0.004794083 39.85841 -104.66700 32.89595  -97.03720
4 LAX SFO 139427 0.006360250 33.94254 -118.40807 37.61900 -122.37484
5 DFW IAH 137004 0.007899200 32.89595  -97.03720 29.98047  -95.33972
6 DTW ORD 135772 0.009424311 42.21206  -83.34884 41.97960  -87.90446

```

Using the above data, the following figure draws lines from ORD (Chicago) to other destinations. 
The black points are the airports that handle 90% of the total air traffic volume. The grey points 
are the remaining airports. The flights from Chicago (ORD) are color coded based on volume carried 
e.g. red implies those routes carry the top 25% of traffic in/out of ORD.

