## Transforming Text Data ##

This chapter builds on the Airline Dataset. One foreseeable use of RHIPE is to transform text 
data. For example:

1. Subset Southwest Airline and Delta Airways information to create a new set of text files, 
one with only Southwest and the other with Delta.

2. Transform the original text data to one with fewer columns and some transformed e.g. 
Airport Codes to full names.

We’ll cover both examples.

### Subset ###

Recall that the airline data take the form
```
1995,1,6,5,657,645,952,937,UA,482,N7298U,115,112,83,15,12,ORD,PHL,678,7,25,0,NA,0,NA,NA,NA,NA,NA
1995,1,7,6,648,645,938,937,UA,482,N7449U,110,112,88,1,3,ORD,PHL,678,5,17,0,NA,0,NA,NA,NA,NA,NA
...
```
The carrier name is column 9. The carrier code for Southwest Airlines is WN, and the code for 
Delta Airways is DL. Only those rows with column 9 equal to WN or DL will be saved.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  tkn  <- strsplit(unlist(map.values), ",")
  text <- do.call("rbind", tkn)
  text <- text[text[, 9] %in% c("WN", "DL"), , drop = FALSE]
  if (nrow(text) > 0)
    apply(text, 1, function(r) rhcollect(r[9], r))
})
```

Note that ```rhcollect()``` requires both a key and value. However, since the key is not used, 
```NULL``` is passed to the key argument and ```mapred.textoutputformat.usekey``` is set to 
```FALSE``` so that the key is not written to disk. By default RHIPE includes strings in 
quotes, but since we do not wish to do so here, we set ```rhipe_string_quote``` to ```''``` 
and ```mapred.field.separator``` to ```","``` as the original data is comma separated. A 
partitioner is used to send all the flight information belonging to Southwest Airlines to one 
file and that belonging to Delta Airways to another.

```{r eval=FALSE, tidy=FALSE}
z <- rhwatch(
  map      = map,
  reduce   = rhoptions()$templates$identity,
  input    = rhfmt("/tmp/airline/data/outputs/1995.csv", type = 'text'),
  output   = rhfmt("/tmp/airline/output/transform", type = 'text'),
  orderby  = "char",
  part     = list(lims = 1, type = "string"),
  mapred   = list(
    mapred.reduce.tasks            = 2,
    rhipe_string_quote             = '',
    mapred.field.separator         = ",",
    mapred.textoutputformat.usekey = FALSE),
  readback = FALSE
  )
```

The output, in one file is
```
1995,11,15,3,937,930,1016,1011,DL,2016,N319DL,39,41,33,5,7,FAT,RNO,188,5,1,0,NA,0,NA,NA,NA,NA,NA
1995,11,16,4,927,930,1017,1011,DL,2016,N326DL,50,41,38,6,-3,FAT,RNO,188,4,8,0,NA,0,NA,NA,NA,NA,NA
1995,11,17,5,931,930,1016,1011,DL,2016,N331DL,45,41,31,5,1,FAT,RNO,188,5,9,0,NA,0,NA,NA,NA,NA,NA
1995,11,18,6,929,930,1012,1011,DL,2016,N237WA,43,41,32,1,-1,FAT,RNO,188,6,5,0,NA,0,NA,NA,NA,NA,NA
1995,11,19,7,928,930,1008,1011,DL,2016,N318DL,40,41,31,-3,-2,FAT,RNO,188,4,5,0,NA,0,NA,NA,NA,NA,NA
```
and in the other is
```
1995,5,1,1,1706,1700,1750,1740,WN,1228,N105,164,160,154,10,6,MCI,PHX,1044,4,6,0,NA,0,NA,NA,NA,NA,NA
1995,4,1,6,630,630,825,825,WN,308,N83,55,55,43,0,0,LAS,PHX,256,2,10,0,NA,0,NA,NA,NA,NA,NA
1995,4,3,1,630,630,719,725,WN,308,N386,49,55,38,-6,0,LAS,PHX,256,3,8,0,NA,0,NA,NA,NA,NA,NA
1995,4,4,2,630,630,720,725,WN,308,N27,50,55,42,-5,0,LAS,PHX,256,2,6,0,NA,0,NA,NA,NA,NA,NA
1995,4,5,3,630,630,723,725,WN,308,N82,53,55,41,-2,0,LAS,PHX,256,3,9,0,NA,0,NA,NA,NA,NA,NA
```

### Transform ###

In this example, we convert each airport code to their name equivalent. Airport codes can be 
found at the [JSM website](http://stat-computing.org/dataexpo/2009/the-data.html). When 
working with massive data, repeatedly used operations need to be as fast as possible. Thus we 
will save the airport code to airport name as a hash table using the ```new.env()``` function. 
Airport codes (origin and destination) are in columns 17 and 18. The setup expression loads 
this data set and creates a function that does the mapping.

```{r eval=FALSE, tidy=FALSE}
airport <- read.table("~/tmp/airports.csv",
                      sep = ",", header = TRUE, stringsAsFactors = FALSE)
aton <- new.env()
for (i in 1:nrow(airport)) {
  aton[[ airport[i, "iata"] ]] <-
    list(ap = airport[i, "airport"], latlong = airport[i, c("lat", "long")])
}
rhsave(aton, file = "/tmp/airline/airline.names/airports.RData")

setup <- expression(
  map = {
    load("airports.RData")
    co <- function(N) {
      sapply(text[, N], function(r) {
        o <- aton[[ r[1] ]]$ap
        if (is.null(o)) NA else sprintf('%s', o)
        }
      })
    }
  })
```

The map function will use the `aton` dictionary to get the complete names.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  tkn        <- strsplit(unlist(map.values), ",")
  text       <- do.call("rbind", tkn)
  text[, 17] <- co(17)
  text[, 18] <- co(18)
  apply(text, 1, function(r) { rhcollect(NULL, r) })
})

z <- rhwatch(
  map      = map,
  reduce   = rhoptions()$templates$identity,
  input    = rhfmt("/tmp/airline/data/outputs/1995.csv", type = 'text'),
  output   = rhfmt("/tmp/airline/output/transform", type = 'text'),
  shared   = c("/tmp/airline/airline.names/airports.RData"),
  setup    = setup,
  mapred   = list(
    mapred.reduce.tasks            = 0,
    rhipe_string_quote             = '',
    mapred.field.separator         = ",",
    mapred.textoutputformat.usekey = FALSE),
  readback = FALSE
)
```

and this gives us

```
1995,1,6,5,657,645,952,937,UA,482,N7298U,115,112,83,15,12,Chicago O'Hare International,Philadelphia Intl,678,7,25,0,NA,0,NA,NA,NA,NA,NA
1995,1,7,6,648,645,938,937,UA,482,N7449U,110,112,88,1,3,Chicago O'Hare International,Philadelphia Intl,678,5,17,0,NA,0,NA,NA,NA,NA,NA
1995,1,8,7,649,645,932,937,UA,482,N7453U,103,112,83,-5,4,Chicago O'Hare International,Philadelphia Intl,678,3,17,0,NA,0,NA,NA,NA,NA,NA
1995,1,9,1,645,645,928,937,UA,482,N7288U,103,112,84,-9,0,Chicago O'Hare International,Philadelphia Intl,678,3,16,0,NA,0,NA,NA,NA,NA,NA
1995,1,10,2,645,645,931,937,UA,482,N7275U,106,112,82,-6,0,Chicago O'Hare International,Philadelphia Intl,678,6,18,0,NA,0,NA,NA,NA,NA,NA
```