## Simulations ##

### Simulations are Embarrassingly Parallel ###

Simulations are an example of task parallel routines in which a function is called repeatedly with 
varying parameters. These computations are processor intensive and consume/produce little data. 
The evaluation of these tasks are independent in that there is no communication between them. With 
`N` tasks and `P` processors, if `P = N` we could run all N in parallel and collect the results. 
However, often `P << N` and thus we must either

* Create a queue of tasks and assign the top most task on the queue to the next free processor. 
This works very well in an heterogeneous environment e.g. with varying processor capacities or 
varying task characteristics - free resources will be automatically assigned pending tasks. The 
cost in creating a new task can be much greater than the cost of evaluating the task.
* Partition the `N` tasks into `n` splits each containing $\lceil N/n \rceil$ tasks (with the last 
split containing the remainder). These splits are placed in a queue, each processor is assigned a 
splits and the tasks in a split are evaluated sequentially.

The second approach simplifies to the first when `n = N`. Creating one split per task is 
inefficient since the time to create,assign launch the task contained in a split might be much 
greater than the evaluation of the task. Moreover, with `N` in the millions, this will cause the 
Jobtracker to run out of memory. It is recommended to divide the `N` tasks into fewer splits of 
sequential tasks. Because of non uniform running times among tasks, processors can spend time in 
the sequential execution of tasks in a split $\sigma$ with other processors idle. Hadoop will 
schedule the split $\sigma$ to another processor (however it will not divide the split into smaller 
splits), and the output of whichever completes first will be used.

RHIPE provides two approaches to this sort of computation. To apply the function `F` to the set $\lbrace 1, 2,\ldots, M \rbrace$, the pseudo code would follow as (here we assume `F` returns a data frame)
```{r eval=FALSE, tidy=FALSE}
FC <- expression({
  results <- do.call("rbind", lapply(map.values, F))
  rhcollect(1, results)
})
mrFC <- rhwatch(
  map    = FC,
  input  = c(1000, 8),
  output = "/tmp/FC",
  inout  = c('lapply', 'sequence'),
  mapred = list(mapred.map.tasks = 1000)
  )
do.call('rbind',lapply(rhread('/tempfolder', mc=TRUE),'[[',2))
```
Here `F` is applied to the numbers $1, 2,\ldots, M$. The job is decomposed into `1000` splits 
(specified by `mapred.map.tasks`) each containing approximately $\lceil M/1000 \rceil$ tasks. 
The expression, `FC` sequentially applies `F` to the elements of `map.values` (which will 
contain a subset of $1, 2,\ldots ,M$) and aggregate the returned data frames with a call to rbind. 
In the last line, the results of the `1000` tasks (which is a list of data frames) are read from 
the HDFS, the data frame are extracted from the list and combined using a call to rbind. Much of 
this is boiler plate RHIPE code and the only varying portions are: the function `F`, the number of 
iterations `M`, the number of groups (e.g. `mapred.map.tasks`) and the aggregation scheme (e.g. I 
used the call to rbind). R lists can be written to a file on the HDFS (with `rhwrite`), which can 
be used as input to a MapReduce job.

### Medians of Standard Normal Samples Example ###

The following is example code for how to generate random deviates and store the medians of each 
subset to the HDFS. This example will generate a total of `N <- 2^18` standard normal deviates in 
`R <- 2^8` subsets of size `m <- 2^10` and reduce to the median value of each subset:
```{r eval=FALSE, tidy=FALSE}
dir <- "tmp/rnorm/"
N <- 2^18
m <- 2^10
mapZ <- expression({
  m <- 2^10
  for(i in seq_along(map.values)){
    Z <- rnorm(m)
    med <- median(Z)
    rhcollect(NULL, med)
  }
})
reduceZ <- expression({rhcollect(reduce.key, reduce.values)})
mrZ <- rhwatch(
  map      = mapZ,
  reduce   = reduceZ,
  input    = c(N/m, 8), 
  output   = paste(dir, "Z", sep=""),
  mapred   = list(mapred.reduce.tasks = 1),
  readback = FALSE
  )
```
Observe the use of `NULL` in the key component of `rhcollect()` within the `mapZ` expression(). 
This choice is, because there is no need to differentiate one simulation from another as being 
unique or special. This makes it easier to work with this particular simulation when bringing this 
data into your R Global Environment.

### Observing Location of Median HDFS Data ###

A simple call to list the available files within your HDFS directory through R utilizing the RHIPE 
library command `rhls()` will display where the random values are stored.
```{r eval=FALSE, tidy=FALSE}
rhls("/tmp/rnorm/Z")
```
```
    permission   owner      group    size          modtime
1   -rwr---r-- jtroisi supergroup       0 2014-06-30 13:53
2   drwxrwxrwt jtroisi supergroup       0 2014-06-30 13:53
3   -rw-r--r-- jtroisi supergroup 6.15 kb 2014-06-30 13:53
                         file
1       /tmp/rnorm/Z/_SUCCESS
2          /tmp/rnorm/Z/_logs
3   /tmp/rnorm/Z/part-r-00000
```
Files are list as above.

### Reading Median HDFS Data into Global Environment ###

To call any of these sets of random numbers into your local environment it is as simple as a call 
to the RHIPE command `rhread()`.
```{r eval=FALSE, tidy=FALSE}
Zmedians <- unlist(rhread("/tmp/rnorm/Z/part-r-00000"))
head(Zmedians)
```
```{r echo=FALSE}
Zmedians <- 1:6
for(i in 1:6) Zmedians[i] <- median(rnorm(2^10))
Zmedians
```
### Nonparametric Spectrum Estimation ###

Averaging Transformation Smoothing (ATS) Methods is a general nonparametric approach that can be utilized on a multitude of data forms. RHIPE is the optimum platform to utilize ATS methods for spectrum estimation.

Below is a toy example, seeing as it is estimating AR(2) simulated data, which would not require such advanced methods. However, the code is functional and would be utilized to provide a nonparametric estimate to univariate time series data of any unknown form.
```{r eval=FALSE, tidy=FALSE}
N <- 2^20
m <- 2^15
rho.true <- c(2/3, -1/3)
mapA <- expression({
  m <- 2^15
  rho.true <- c(2/3, -1/3)
  for(i in seq_along(map.values)){
    AR2 <- arima.sim(n = m, model = list(ar = rho.true))
    X <- fft(AR2)
    X <- X[2:(m/2)]
    per <- Mod(X)^2
    rhcollect(NULL, per)
  }
})
reduceA <- expression(
  pre    = {A <- rep(0, m/2 - 1)},
  reduce = {A <- A + apply(do.call("rbind", reduce.values), 2, sum)},
  post   = {rhcollect(NULL, A)}
  )
mrA <- rhwatch(
  map      = mapA,
  reduce   = reduceA,
  input    = c(N/m, 8),
  output   = paste(dir, "A", sep=""),
  mapred   = list(mapred.reduce.tasks = 1),
  readback = FALSE
  )
mapT <- expression({
  for(i in seq_along(map.values)){
    A <- map.values[[i]]
    T <- log(A)
    rhcollect(NULL, T)
  }
})
mrT <- rhwatch(
  map = mapT,
  input = paste(dir, "A", sep=""),
  output = paste(dir, "T", sep=""),
  mapred = list(mapred.reduce.tasks = 0),
  readback = FALSE
  )
```

### Plotting LOESS in R Global Environment ###

```{r eval=FALSE, tidy=FALSE}
dirT <- "/ln/jtroisi/T/part-m-00000"
T <- unlist(rhread(dirT))
freq <- time(T)/(m + 1)
library(lattice)
trellis.device(postscript, file = "S.ps")
xyplot(T ~ freq,
       panel = function(x, y){
         panel.loess(x, y, span = 101/(m + 1), degree = 2, col = "red")
         panel.xyplot(x, y, pch = ".")
       }
       )
dev.off()
```
Another useful plot would be a residuals plot, but it is not necessary to demonstrate the R coding for such a task here.